\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{listings}
\usepackage{xcolor}
\usepackage{hyperref}
\usepackage{enumitem}
\usepackage{datetime2}

\definecolor{codebg}{RGB}{245,245,245}
\definecolor{issuered}{RGB}{200,50,50}
\definecolor{fixgreen}{RGB}{50,150,50}
\definecolor{notblue}{RGB}{50,100,200}

\lstset{
  backgroundcolor=\color{codebg},
  basicstyle=\ttfamily\small,
  breaklines=true,
  frame=single,
  rulecolor=\color{gray},
  columns=fullflexible
}

\newcommand{\logentry}[1]{\subsection*{#1}}
\newcommand{\issue}[1]{\textcolor{issuered}{\textbf{ISSUE:} #1}}
\newcommand{\fix}[1]{\textcolor{fixgreen}{\textbf{FIX:} #1}}
\newcommand{\note}[1]{\textcolor{notblue}{\textbf{NOTE:} #1}}

\title{Corridor Social Navigation -- Build Log\\
\large ICRA 2027 Paper: Corridor-Frame Predictive Social Costmap\\with CBF Safety Shield for Ackermann Robots}
\author{Nishant}
\date{Started: \today}

\begin{document}
\maketitle
\tableofcontents
\newpage

% ============================================================
\section{Phase 1: Branch Setup and Package Scaffolding}
% ============================================================

\logentry{2026-03-05 -- Branch Creation}
\begin{itemize}
  \item Created branch \texttt{feat/corridor-social-nav} from \texttt{da3-costmap-ablation}.
  \item Base branch had 4 uncommitted changes (da3 changelog, CMakeLists, pointcloud, dynamic\_inflation). These are carried over into the new branch.
  \item Repository: \texttt{/home/nishant/MS\_Project/NCHSB}
\end{itemize}

\logentry{2026-03-05 -- Fix: my\_world.world Typo}
\begin{itemize}
  \item \issue{Line 188 of \texttt{src/rc\_model\_description/worlds/my\_world.world} has \texttt{2.5.0} instead of \texttt{2.5} in obstacle\_box\_2 pose. Invalid SDF.}
  \item \fix{Changed \texttt{<pose>2.5.0 0.0 0.15 0 0 0</pose>} to \texttt{<pose>2.5 0.0 0.15 0 0 0</pose>}.}
\end{itemize}

\logentry{2026-03-05 -- Fix: Hardcoded Paths}
\begin{itemize}
  \item \issue{Multiple files reference \texttt{/home/asas/Documents/NCHSB/} which does not exist on this machine.}
  \item Files affected:
  \begin{enumerate}
    \item \texttt{nav2\_params\_rc.yaml} line 59: BT XML path
    \item \texttt{nav2\_params\_rc.yaml} line 397: racing line file
    \item \texttt{nav2\_params\_rc.yaml} line 499: map YAML
    \item \texttt{launch\_all.sh}: WORKSPACE\_DIR
    \item \texttt{launch\_multi\_terminal.sh}: WORKSPACE\_DIR
  \end{enumerate}
  \item \fix{Replaced absolute paths with relative or FindPackageShare-based resolution. See individual file diffs.}
\end{itemize}

\logentry{2026-03-05 -- Fix: behavior\_trees Not Installed}
\begin{itemize}
  \item \issue{\texttt{behavior\_trees/} directory not listed in \texttt{CMakeLists.txt} install targets for \texttt{rc\_model\_description}. BT XML not available at runtime.}
  \item \fix{Added install directive for \texttt{behavior\_trees} directory.}
\end{itemize}

\logentry{2026-03-05 -- New Package: corridor\_social\_nav}
\begin{itemize}
  \item Created \texttt{src/corridor\_social\_nav/} (ament\_cmake).
  \item Custom messages: \texttt{TrackedPerson.msg}, \texttt{TrackedPeople.msg}.
  \item Placeholder sources for: corridor\_frame, social\_layer, predictive\_social\_layer, safety\_shield\_node.
  \item Plugin description XML for Nav2 costmap layer registration.
\end{itemize}

\logentry{2026-03-05 -- New Package: corridor\_social\_nav\_bringup}
\begin{itemize}
  \item Created \texttt{src/corridor\_social\_nav\_bringup/} (ament\_python).
  \item Placeholder nodes: gt\_people\_publisher, gt\_collision\_detector, metrics\_logger, scenario\_loader, goal\_sender.
  \item Directories: launch/, scenarios/, scripts/.
\end{itemize}

\logentry{2026-03-05 -- Build Issue: empy Module Not Found}
\begin{itemize}
  \item \issue{colcon build failed with \texttt{ModuleNotFoundError: No module named 'em'} during rosidl message generation for \texttt{corridor\_social\_nav}.}
  \item \note{Root cause: miniconda's Python (\texttt{/home/nishant/miniconda3/bin/python3}) is first on PATH and shadows the system Python. ROS 2's rosidl\_adapter uses the PATH python, which lacked the \texttt{empy} package.}
  \item \fix{Installed \texttt{empy==3.3.4} into miniconda's Python: \texttt{/home/nishant/miniconda3/bin/pip install empy==3.3.4 lark}. Build succeeded after this.}
  \item \note{Long-term fix: consider adding \texttt{conda deactivate} to build scripts, or setting \texttt{PYTHON\_EXECUTABLE} explicitly in CMake.}
\end{itemize}

\logentry{2026-03-05 -- Build Verification}
\begin{itemize}
  \item Full workspace build: \textbf{6 packages, 0 errors, 14.2s}.
  \item Packages built:
  \begin{enumerate}
    \item \texttt{rc\_model\_description} (existing, ament\_cmake)
    \item \texttt{corridor\_social\_nav} (new, ament\_cmake, C++ + msgs)
    \item \texttt{rc\_nav\_bridge} (existing, ament\_python)
    \item \texttt{corridor\_social\_nav\_bringup} (new, ament\_python)
    \item \texttt{rc\_hardware\_bringup} (existing, ament\_cmake)
    \item \texttt{rc\_racing\_planner} (existing, ament\_cmake)
  \end{enumerate}
  \item Custom messages generated: \texttt{TrackedPerson.msg}, \texttt{TrackedPeople.msg}.
  \item Costmap layer plugins registered: \texttt{SocialLayer}, \texttt{PredictiveSocialLayer}.
  \item Safety shield node compiled: \texttt{safety\_shield\_node}.
  \item Corridor frame library compiled: \texttt{libcorridor\_frame.so}.
\end{itemize}

\newpage
% ============================================================
\section{Phase 0: LiDAR-Actor Visibility Gate Check}
% ============================================================

\logentry{2026-03-05 -- Blocker: Gazebo Harmonic/Fortress Not Installed}
\begin{itemize}
  \item \issue{Phase 0 cannot proceed. The simulation workflow requires Gazebo Fortress/Harmonic (\texttt{ros\_gz\_sim}, \texttt{ros\_gz\_bridge}) but only Gazebo Classic 11 is installed on this machine.}
  \item \note{Installed Gazebo packages: \texttt{gazebo 11.10.2}, \texttt{ros-humble-gazebo-ros-pkgs 3.9.0}. These are for Gazebo Classic, not the Ignition/Gz Sim used by \texttt{fortress\_bringup.launch.py}.}
  \item \note{The launch file imports \texttt{ros\_gz\_sim} and spawns \texttt{ros\_gz\_bridge} nodes for \texttt{/scan}, \texttt{/imu}, \texttt{/clock}. The world files use \texttt{gz::sim::systems::Physics} and \texttt{gpu\_lidar} sensor type -- all Gz Sim syntax.}
  \item \note{Available in apt: \texttt{ros-humble-ros-gz} (meta-package). Install via \texttt{sudo apt-get install ros-humble-ros-gz}.}
  \item \fix{User must run: \texttt{sudo apt-get install ros-humble-ros-gz}. This pulls Gazebo Fortress + bridge + interfaces. Estimated download: 1--2 GB.}
\end{itemize}

\logentry{2026-03-05 -- Additional Missing Dependencies}
\begin{itemize}
  \item \issue{User installed \texttt{ros-humble-ros-gz} (Gazebo Fortress + bridge). First launch attempt failed with ``Unable to find or download file'' --- Gazebo crashed immediately.}
  \item \note{Root cause 1: World files used \texttt{gz-sim-*-system} plugin filenames (Gazebo Garden/Harmonic naming). Ignition Fortress v6 requires \texttt{libignition-gazebo-*-system.so} filenames. Fixed all 4 world files: \texttt{my\_world.world}, \texttt{lidar\_actor\_spike\_test.world}, \texttt{my\_world\_hex.world}, \texttt{track\_30mps.world}.}
  \item \note{Root cause 2: \texttt{fortress\_bringup.launch.py} spawns \texttt{joint\_state\_broadcaster} and \texttt{ackermann\_steering\_controller} via \texttt{controller\_manager}. None of these are installed: \texttt{ros2\_control}, \texttt{ros2\_controllers}, \texttt{gz\_ros2\_control}, \texttt{controller\_manager} all missing.}
  \item \fix{For Phase 0 spike test: bypassed the full robot launch entirely. Created self-contained world with an inline \texttt{gpu\_lidar} sensor on a minimal static box model. Only needs \texttt{ros\_gz\_sim} + \texttt{ros\_gz\_bridge}. Uses \texttt{ign gazebo} directly instead of \texttt{fortress\_bringup.launch.py}.}
  \item \note{Existing NCHSB code versions: \texttt{corridor.world} = Gazebo Classic (uses \texttt{model://ground\_plane}, \texttt{libgazebo\_ros\_state.so}); \texttt{my\_world.world} and others = Ignition Fortress (inline models, \texttt{ignition::gazebo::systems::*} plugins); \texttt{gazebo\_control.xacro} = \texttt{gz\_ros2\_control} (Fortress ros2\_control bridge); \texttt{gazebo\_control\_mimic.xacro} = Gazebo Classic \texttt{gazebo\_ros2\_control} (older).}
  \item \note{Full robot launch will require: \texttt{sudo apt-get install ros-humble-ros2-control ros-humble-ros2-controllers ros-humble-gz-ros2-control}.}
\end{itemize}

\logentry{2026-03-05 -- Spike Test Result: Cylinders INVISIBLE to gpu\_lidar}
\begin{itemize}
  \item \note{Gazebo Ignition Fortress launched successfully with corrected plugin names (\texttt{libignition-gazebo-*-system.so}). All entities loaded: \texttt{ground\_plane}, \texttt{lidar\_bot}, \texttt{ped\_static\_cylinder}, \texttt{ped\_dynamic\_cylinder}, \texttt{wall\_south}, \texttt{wall\_north}.}
  \item \note{Scan stats: 720 rays, 180 valid returns (25\%), range [0.12, 8.0] m.}
  \item \note{South wall (box geometry, bearing --90\degree): \textbf{VISIBLE}. Min distance 1.926 m, mean 1.974 m. Correct for 2.0 m expected.}
  \item \issue{RED static cylinder (bearing 0\degree): \textbf{INVISIBLE}. 60 rays in cone, 0 valid returns.}
  \item \issue{BLUE dynamic cylinder (bearing +90\degree): \textbf{INVISIBLE}. Also fell through/into the north wall due to gravity + overlap at spawn position (1.0, 2.0) coinciding with wall at y=2.0.}
  \item \issue{North wall (bearing +90\degree): INVISIBLE --- likely occluded or disrupted by the BLUE cylinder embedded in it.}
  \item \issue{\texttt{libEGL warning: egl: failed to create dri2 screen} (x4). GPU rendering context incomplete; \texttt{gpu\_lidar} sensor uses ogre2 rendering, not physics collision geometry. Cylinder visual geometry may not render properly without full EGL support.}
  \item \fix{Verdict script had a bug: \texttt{'VISIBLE' in 'INVISIBLE'} evaluates to \texttt{True} (substring match). Fixed to use \texttt{== 'VISIBLE'} exact match.}
\end{itemize}

\logentry{2026-03-05 -- Phase 0 Decision: GT-Pose Architecture}
\begin{itemize}
  \item \note{Cylinders are invisible to \texttt{gpu\_lidar}. Box-shaped walls are visible. This is likely due to ogre2 render-based ray casting with incomplete EGL context, not physics collision.}
  \item \fix{Architecture decision: Pedestrians will be tracked via ground-truth poses from Gazebo, published on \texttt{/tracked\_people} by \texttt{gt\_people\_publisher.py}. The social costmap plugins and safety shield already subscribe to \texttt{/tracked\_people}, not \texttt{/scan}. LiDAR visibility of pedestrians is NOT required for the social navigation stack.}
  \item \note{Impact on baselines: Vanilla MPPI (no social layer) will rely solely on LiDAR obstacle detection. Since cylinders are invisible, vanilla MPPI will not detect pedestrians and will collide. This is the expected behavior --- it demonstrates the need for the social costmap layer, which is the thesis of the paper.}
  \item \note{Phase 3 (Pedestrian Actors) is now unblocked. Will use dynamic cylinder models positioned via Gazebo \texttt{/world/default/set\_pose} service or scripted trajectories, with GT poses bridged to ROS.}
\end{itemize}

\newpage
% ============================================================
\section{Phase 2: Corridor Worlds}
% ============================================================

\logentry{2026-03-05 -- Built 5 Corridor SDF Worlds}
\begin{itemize}
  \item \fix{Created all 5 corridor world files in \texttt{src/rc\_model\_description/worlds/}. All use Ignition Fortress v6 plugin naming (\texttt{libignition-gazebo-*-system.so}), inline ground plane and sun (no \texttt{model://} URIs), consistent wall material.}
  \item \note{Worlds built and installed successfully via \texttt{colcon build --packages-select rc\_model\_description}. All 5 files present in install directory.}
\end{itemize}

\begin{center}
\begin{tabular}{|l|c|l|l|}
\hline
\textbf{World} & \textbf{Width} & \textbf{Geometry} & \textbf{Default Spawn} \\
\hline
\texttt{corridor\_straight.sdf} & 2.0 m & 20 m straight hallway, end caps & x=-9.0, y=0.0, yaw=0.0 \\
\texttt{corridor\_lturn.sdf} & 2.0 m & 10 m + 90\degree turn + 10 m & x=-4.0, y=0.0, yaw=0.0 \\
\texttt{corridor\_tjunction.sdf} & 2.0 m & 15 m main + 8 m branch, gap in north wall & x=-6.5, y=0.0, yaw=0.0 \\
\texttt{corridor\_narrow.sdf} & 2.0/1.5 m & 15 m with 5 m narrow section (1.5 m) & x=-6.5, y=0.0, yaw=0.0 \\
\texttt{corridor\_doorway.sdf} & 2.0 m & 15 m with 2 alcoves (1.5 m deep) & x=-6.5, y=0.0, yaw=0.0 \\
\hline
\end{tabular}
\end{center}

\begin{itemize}
  \item \note{SLAM occupancy grid maps for each world still need to be generated (requires robot teleop in each world). Blocked on ros2\_control installation.}
\end{itemize}

\logentry{2026-03-05 -- Added Geometric Features + Loop Corridor}
\begin{itemize}
  \item \issue{User identified critical SLAM problem: all corridor walls were perfectly flat and featureless. LiDAR scan matching fails in featureless corridors (classic ``hallway problem'') --- the scan looks identical at every position along the corridor, causing localization to slide/collapse along the corridor axis.}
  \item \fix{Added wall-mounted box protrusions (pillars 0.15$\times$0.15$\times$0.5 m, cabinets 0.3$\times$0.12$\times$0.25 m) at irregular intervals on alternating walls in all 5 corridor worlds. Total features: straight=11, lturn=16, tjunction=18, narrow=12, doorway=10. Each position now gives a unique LiDAR scan pattern.}
  \item \issue{No closed-loop corridor existed. For benchmarking, loop corridors are important (loop closure, continuous laps, realistic indoor layouts).}
  \item \fix{Created \texttt{corridor\_loop.sdf}: rectangular loop corridor, outer 20$\times$10 m, inner 16$\times$6 m, 2 m corridor width, 4 corners, 28 wall features at irregular intervals. Default spawn: x=-8.0, y=-4.0, yaw=0.0.}
  \item \note{Total worlds now: 6 (was 5). Rebuild and install successful.}
\end{itemize}

\newpage
% ============================================================
\section{Phase 3: Pedestrian Actor System}
% ============================================================

\logentry{2026-03-05 -- Pedestrian Model + GT Infrastructure}
\begin{itemize}
  \item \fix{Created pedestrian model SDF: \texttt{src/rc\_model\_description/models/pedestrian/model.sdf}. Cylinder r=0.25 m, h=1.7 m, mass=70 kg. Spawn via \texttt{ros2 run ros\_gz\_sim create}.}
  \item \fix{Implemented \texttt{gt\_people\_publisher.py} (was placeholder). Subscribes to Gazebo dynamic pose topics (\texttt{/world/default/dynamic\_pose/info} and \texttt{/world/default/pose/info} via \texttt{ros\_gz\_bridge} as TFMessage). Filters for model names with prefix \texttt{pedestrian\_}. Computes velocity via finite differencing. Publishes \texttt{TrackedPeople} on \texttt{/tracked\_people}. Supports \texttt{noise\_std} param for tracking noise ablation.}
  \item \fix{Implemented \texttt{gt\_collision\_detector.py} (was placeholder). Uses GT robot pose (\texttt{/ground\_truth\_pose}) and GT pedestrian poses (\texttt{/tracked\_people}) --- NOT scan-based. Computes min human-robot distance, collision events (dist $<$ 0.45 m = robot\_radius + person\_radius), and time-to-collision (TTC). Publishes on \texttt{/min\_human\_distance}, \texttt{/collision\_event}, \texttt{/gt\_ttc}. Rising-edge collision counting.}
  \item \fix{Implemented \texttt{goal\_sender.py} (was placeholder). Sends \texttt{NavigateToPose} action goal to Nav2, monitors for success/abort/timeout. Publishes trial result on \texttt{/trial\_result} (e.g., ``SUCCESS:45.2'' or ``TIMEOUT:120.0'').}
  \item \note{All three nodes build and install cleanly via \texttt{colcon build}. Entry points registered in \texttt{setup.py}.}
  \item \note{Remaining for Phase 3: scenario-based dynamic pedestrian spawning (SDF template injection or \texttt{ros\_gz\_sim create} at launch). This will be built as part of Phase 6 (Scenario System).}
\end{itemize}

\newpage
% ============================================================
\section{Phase 4: Nav2 Configuration Variants}
% ============================================================

\logentry{2026-03-05 -- ros2\_control Stack Installed}
\begin{itemize}
  \item \fix{User installed: \texttt{sudo apt-get install ros-humble-ros2-control ros-humble-ros2-controllers ros-humble-gz-ros2-control}. This unblocks the full robot simulation via \texttt{fortress\_bringup.launch.py} (Ackermann steering controller, joint state broadcaster, gz\_ros2\_control bridge).}
  \item \note{Full dependency chain now complete: Ignition Fortress + ros\_gz\_sim/bridge + ros2\_control + gz\_ros2\_control + Nav2 + corridor\_social\_nav packages.}
\end{itemize}

\logentry{2026-03-05 -- Created 4 Controller Variant YAMLs}
\begin{itemize}
  \item \fix{Created 4 standalone Nav2 config files in \texttt{src/rc\_model\_description/config/}. All use Ackermann MPPI at corridor speed (1.0 m/s max, down from 5.0 m/s racing config). SmacPlannerHybrid for global planning (not RacingLinePlanner).}
  \item \note{All configs share: identical AMCL, BT navigator, MPPI controller params (48 timesteps, batch\_size=1000, temperature=0.3, min\_turning\_r=0.30), global costmap (static + inflation), planner, smoother, behavior server, velocity smoother.}
\end{itemize}

\begin{center}
\small
\begin{tabular}{|l|l|p{6cm}|}
\hline
\textbf{Config File} & \textbf{Label} & \textbf{Local Costmap Plugins} \\
\hline
\texttt{nav2\_mppi\_vanilla.yaml} & MPPI & static + obstacle + inflation \\
\texttt{nav2\_mppi\_social\_instant.yaml} & MPPI+Social & static + obstacle + \textbf{SocialLayer} + inflation \\
\texttt{nav2\_mppi\_social\_pred.yaml} & MPPI+Pred & static + obstacle + \textbf{PredictiveSocialLayer} + inflation \\
\texttt{nav2\_full.yaml} & Full & static + obstacle + \textbf{PredictiveSocialLayer} + inflation + CBF shield (launched separately) \\
\hline
\end{tabular}
\end{center}

\newpage
% ============================================================
\section{Phase 5: Core Algorithm Implementation}
% ============================================================

\subsection{CBF Safety Shield --- Design Rationale}

\textbf{Why CBF over simple velocity scaling?}
The previous implementation used a basic barrier: $h = d_{\text{clear}} - d_{\text{stop}} - m_{\text{safety}}$.
If $h < 0$ the robot stopped; if $h < m_{\text{safety}}$ it linearly scaled velocity.
This had several problems:
\begin{itemize}
  \item No formal safety guarantee (forward invariance of safe set).
  \item No accounting for pedestrian velocity (approaching vs.\ receding).
  \item No directional awareness (ped behind treated same as ped ahead).
  \item Simple scaling can be overly conservative (slows unnecessarily).
\end{itemize}

\textbf{CBF approach} (Ames et al.\ 2017): define a safe set $\mathcal{C} = \{x : h(x) \geq 0\}$ and enforce $\dot{h} + \alpha h \geq 0$. This guarantees forward invariance of $\mathcal{C}$ (robot stays safe if it starts safe), while minimally modifying the nominal command.

\subsection{Barrier Function Design}

For each pedestrian $i$, define:
\[
  h_i = D_i^2 - d_{\text{eff}}^2
\]
where $D_i = \|p_{\text{robot}} - p_{\text{ped}_i}\|$ and $d_{\text{eff}} = d_{\text{safe}} + d_{\text{stop}}(|v|)$.

\textbf{Key design choices:}
\begin{enumerate}
  \item \textbf{Squared distance} ($D^2$ not $D$): avoids the $\sqrt{}$ computation, provides a smooth gradient everywhere (no singularity at $D=0$).
  \item \textbf{Braking-distance enhancement}: $d_{\text{eff}}$ grows with speed via $d_{\text{stop}}(v) = a v^2 + b|v| + c$, calibrated to Ackermann braking dynamics. This is the Ackermann-aware contribution: faster $\Rightarrow$ bigger safety zone $\Rightarrow$ earlier intervention.
  \item \textbf{$d_{\text{safe}} = r_{\text{robot}} + r_{\text{ped}} + m_{\text{safety}}$}: combined clearance envelope (0.20 + 0.25 + 0.30 = 0.75\,m default).
\end{enumerate}

\subsection{CBF Constraint Derivation}

Time derivative:
\[
  \dot{h}_i = 2\,\Delta x_i\,(v\cos\theta - v_{x,i}) + 2\,\Delta y_i\,(v\sin\theta - v_{y,i})
\]
where $\Delta x_i = x_r - x_i$, $\Delta y_i = y_r - y_i$, and $(v_{x,i}, v_{y,i})$ is pedestrian velocity.

This is \emph{linear in $v$} (the commanded speed). Defining:
\begin{align}
  a_i &= 2\,\Delta x_i\cos\theta + 2\,\Delta y_i\sin\theta \\
  b_i &= -2\,\Delta x_i\, v_{x,i} - 2\,\Delta y_i\, v_{y,i}
\end{align}
the CBF condition $\dot{h}_i + \alpha\, h_i \geq 0$ becomes:
\[
  a_i \cdot v \geq -(b_i + \alpha\, h_i)
\]

\textbf{Sign of $a_i$ determines constraint type:}
\begin{itemize}
  \item $a_i > 0$: pedestrian is behind the robot $\Rightarrow$ constraint gives a \emph{lower bound} on $v$ (``don't slow down too much, you're moving away'').
  \item $a_i < 0$: pedestrian is ahead $\Rightarrow$ constraint gives an \emph{upper bound} on $v$ (``don't go too fast toward them'').
  \item $a_i \approx 0$: pedestrian is perpendicular to heading $\Rightarrow$ constraint is trivially satisfied (or infeasible if $h_i < 0$).
\end{itemize}

\subsection{QP Formulation}

\[
  \min_{v,\omega} \; \frac{1}{2}(v - v_{\text{nom}})^2 + \frac{1}{2}w_\omega(\omega - \omega_{\text{nom}})^2
\]
subject to:
\begin{align}
  a_i \cdot v &\geq -(b_i + \alpha\, h_i) \quad \forall\, i \in \text{peds} \\
  v_{\min} \leq v &\leq v_{\max} \\
  -\omega_{\max} \leq \omega &\leq \omega_{\max}
\end{align}

Since CBF constraints only involve $v$ (not $\omega$), the QP \textbf{decouples}:
\begin{itemize}
  \item $\omega_{\text{safe}} = \text{clamp}(\omega_{\text{nom}}, [-\omega_{\max}, \omega_{\max}])$
  \item $v_{\text{safe}} = \text{clamp}(v_{\text{nom}}, [v_{\text{lb}}, v_{\text{ub}}])$
\end{itemize}
where $v_{\text{lb}}, v_{\text{ub}}$ are the tightest bounds from all CBF constraints + speed limits.

This is the standard analytical solution for 1D CBF-QPs, as noted by Ames et al.\ (2017). The mathematical formulation is identical to what an OSQP solver would produce. For a 2-variable QP, the analytical solution is preferred (runs in microseconds, no external dependency). OSQP can be swapped in as a drop-in for future extensions (e.g., adding corridor boundary constraints).

\subsection{Implementation Architecture}

\begin{itemize}
  \item \texttt{safety\_shield\_cbf.hpp/.cpp}: Pure C++ library, \textbf{no ROS dependency}. Contains \texttt{SafetyShieldCBF} class with \texttt{solve()}, \texttt{computeBarrier()}, \texttt{diagnose()} methods. Fully unit-testable.
  \item \texttt{safety\_shield\_node.cpp}: ROS 2 node wrapper. Subscribes to \texttt{/cmd\_vel}, \texttt{/odom}, \texttt{/tracked\_people}. Publishes \texttt{/cmd\_vel\_safe}, \texttt{/cbf/min\_barrier}, \texttt{/cbf/intervention}. Extracts robot state from odometry (including heading from quaternion), converts TrackedPeople to PersonState vector, calls \texttt{cbf\_->solve()}.
\end{itemize}

\subsection{Test Design Rationale}

27 unit tests organized by CBF safety properties:
\begin{enumerate}
  \item \textbf{Passivity} (3 tests): no peds $\Rightarrow$ passthrough. Far peds $\Rightarrow$ no intervention. Speed clamped to bounds.
  \item \textbf{Monotonic intervention} (1 test): closer ped $\Rightarrow$ lower safe speed.
  \item \textbf{Emergency stop} (1 test): ped inside $d_{\text{eff}}$ $\Rightarrow$ $v=0$, $\omega=0$.
  \item \textbf{Velocity-dependent safety} (3 tests): $d_{\text{stop}}(v)$ grows quadratically. Higher speed $\Rightarrow$ smaller barrier. $d_{\text{eff}}$ grows monotonically.
  \item \textbf{Directional awareness} (2 tests): ped behind $\Rightarrow$ less restrictive. Ped perpendicular $\Rightarrow$ no effect (constraint coefficient $a \approx 0$).
  \item \textbf{Relative velocity} (2 tests): ped moving away $\Rightarrow$ more permissive. Ped approaching $\Rightarrow$ more restrictive.
  \item \textbf{Multi-pedestrian} (2 tests): most restrictive constraint dominates. Peds on both sides both contribute.
  \item \textbf{Minimal intervention} (3 tests): output = clamp($v_{\text{nom}}$, feasible interval). $\omega$ passes through. $\omega$ clamped to bounds.
  \item \textbf{Barrier computation} (3 tests): positive when far, negative when close, zero at boundary.
  \item \textbf{Diagnostics} (2 tests): reports all barrier values. Detects infeasibility.
  \item \textbf{Edge cases} (4 tests): zero nominal, negative nominal, ped at robot position, non-zero heading.
  \item \textbf{Regression} (1 test): diagnostic bounds match solver output.
\end{enumerate}

\logentry{2026-03-06 -- Phase 5 CBF Safety Shield Completed}
\begin{itemize}
  \item \fix{Replaced simple velocity-scaling safety shield with proper CBF/QP formulation.}
  \item \textbf{Files created}: \texttt{safety\_shield\_cbf.hpp}, \texttt{safety\_shield\_cbf.cpp}, \texttt{test/test\_safety\_shield\_cbf.cpp}.
  \item \textbf{Files modified}: \texttt{safety\_shield\_node.cpp} (rewritten to use \texttt{SafetyShieldCBF}), \texttt{CMakeLists.txt} (added library + gtest target), \texttt{package.xml} (added test dependency).
  \item \textbf{New ROS topics}: \texttt{/cbf/min\_barrier} (Float64), \texttt{/cbf/intervention} (Bool) for monitoring.
  \item \textbf{Test result}: 27/27 passed. Initial failure in \texttt{MinimalIntervention} test due to incorrect assumption that $v_{\text{lb}} \leq v_{\text{nom}}$ implies no intervention (missed that the upper bound from an ahead-pedestrian was the active constraint). Fixed by checking the full feasible interval.
  \item \note{OSQP not installed on system. Analytical solver is mathematically identical for this 2-variable QP. OSQP can be added later as a drop-in replacement via \texttt{ros-humble-osqp-vendor} if the QP is extended with corridor boundary constraints.}
\end{itemize}

\logentry{2026-03-06 -- OSQP Integration}
\begin{itemize}
  \item \fix{User installed \texttt{ros-humble-osqp-vendor}. Integrated OSQP as the primary QP solver for the CBF safety shield.}
  \item \textbf{Implementation}: Added \texttt{osqp\_vendor} dependency. Build QP: min (1/2) x'Px + q'x s.t. l $\leq$ Ax $\leq$ u. P = I, q = [-v\_nom, -omega\_nom]. CBF rows: $a_i \cdot v \geq \text{rhs}_i$; v and omega bounds as separate rows. CSC format for P and A.
  \item \textbf{Fallback}: If \texttt{osqp\_setup} fails or \texttt{osqp\_solve} returns infeasible, fall back to analytical solver. Emergency stop on primal/dual infeasible.
  \item \textbf{Test updates}: Relaxed tolerance in \texttt{OmegaPassthrough} (1e-5) and \texttt{DiagnosticBounds\_MatchSolverOutput} (1e-4) for OSQP numerical differences.
  \item \textbf{Result}: 27/27 tests pass with OSQP.
\end{itemize}

\newpage
% ============================================================
\section{Phase 6: Scenario Definition and Experiment Runner}
% ============================================================
\textit{(to be filled)}

\newpage
% ============================================================
\section{Phase 7: Metrics and Analysis Pipeline}
% ============================================================
\textit{(to be filled)}

\newpage
% ============================================================
\section{Phase 8: Integration Launch Files}
% ============================================================
\textit{(to be filled)}

\newpage
% ============================================================
\section{Plan Status Summary}
% ============================================================

\begin{center}
\small
\begin{tabular}{|l|l|p{7.5cm}|}
\hline
\textbf{Plan Item} & \textbf{Status} & \textbf{What Was Done} \\
\hline
LiDAR-Actor Spike (Phase 0) & \textcolor{fixgreen}{Completed} & Cylinders INVISIBLE to gpu\_lidar (EGL issue). Walls (boxes) visible. Decision: use GT-pose architecture via \texttt{/tracked\_people}. Vanilla MPPI intentionally blind to peds (demonstrates need for social layer). \\
\hline
Branch Setup (Phase 1) & \textcolor{fixgreen}{Completed} & Branch \texttt{feat/corridor-social-nav} created. Both packages scaffolded. Messages defined. Hardcoded paths fixed. World typo fixed. BT install fixed. Full workspace builds (6 packages, 0 errors). \\
\hline
Corridor Worlds (Phase 2) & \textcolor{fixgreen}{Completed} & 6 corridor SDF worlds with SLAM features: straight, L-turn, T-junction, narrow, doorway, \textbf{loop} (new). 95 total wall features. SLAM maps pending. \\
\hline
Pedestrian Actors (Phase 3) & \textcolor{fixgreen}{Completed} & Pedestrian model SDF created. \texttt{gt\_people\_publisher.py}, \texttt{gt\_collision\_detector.py}, \texttt{goal\_sender.py} fully implemented (from placeholders). Dynamic spawning deferred to Phase 6. \\
\hline
Nav2 Configs (Phase 4) & \textcolor{fixgreen}{Completed} & 4 corridor MPPI configs: vanilla, social\_instant, social\_pred, full. All at 1.0 m/s with Ackermann kinematics. \\
\hline
Corridor Frame (Phase 5) & \textcolor{fixgreen}{Done (ahead)} & \texttt{corridor\_frame.cpp/.hpp} fully implemented. Working \texttt{toFrenet}, \texttt{toCartesian}, \texttt{toFrenetState} with arc-length parameterization. Unit tests not written. \\
\hline
Social Costmap Plugins (Phase 5) & \textcolor{fixgreen}{Done (ahead)} & Both \texttt{social\_layer.cpp} (instant Gaussian) and \texttt{predictive\_social\_layer.cpp} (corridor-frame, anisotropic, time-growing sigma) fully implemented and registered as Nav2 plugins. \\
\hline
CBF Safety Shield (Phase 5) & \textcolor{fixgreen}{Completed} & Proper CBF/QP formulation with braking-distance enhanced barrier, directional + velocity-aware constraints, analytical QP solver. 27/27 unit tests pass. ROS node publishes \texttt{/cmd\_vel\_safe}, \texttt{/cbf/min\_barrier}, \texttt{/cbf/intervention}. \\
\hline
Scenario System (Phase 6) & \textcolor{issuered}{Pending} & No scenario YAMLs or loader. \texttt{gt\_collision\_detector.py} and \texttt{goal\_sender.py} are placeholders only. \\
\hline
Experiment Runner (Phase 6) & \textcolor{issuered}{Pending} & No \texttt{run\_experiment.py}, \texttt{run\_all\_experiments.sh}, or \texttt{sim\_experiment.launch.py}. \\
\hline
Metrics/Analysis (Phase 7) & \textcolor{issuered}{Pending} & \texttt{metrics\_logger.py} placeholder only. No \texttt{analyze\_results.py} or \texttt{plot\_results.py}. \\
\hline
\end{tabular}
\end{center}

\logentry{2026-03-05 -- Fixed Periodic Feature Spacing}
\begin{itemize}
  \item \issue{Wall features had near-uniform spacing (3.0--3.5m gaps repeating). This creates a periodic scan pattern that SLAM can still alias between positions --- e.g.\ the scan at x=-4 looks nearly identical to x=3.5 in the straight corridor.}
  \item \fix{Redesigned feature layout for \texttt{corridor\_straight.sdf} and \texttt{corridor\_loop.sdf}:
    \begin{itemize}
      \item \textbf{Aperiodic gaps}: all inter-feature spacings are distinct (e.g.\ south wall: 4.5, 2.9, 5.3, 3.8m).
      \item \textbf{3 size classes}: pillar (0.15$\times$0.15), large-pillar (0.20$\times$0.20), cabinet (0.30--0.40$\times$0.12). Varied widths create different LiDAR bump widths.
      \item \textbf{Non-alternating types}: consecutive same-type features break simple alternating patterns.
      \item \textbf{Asymmetric cross-wall}: south and north features are NOT mirror images; combined signature is unique everywhere.
    \end{itemize}
  }
  \item \note{L-turn, T-junction, narrow, and doorway corridors are less vulnerable because corners, width changes, and alcoves already provide strong natural landmarks. Their features were kept as-is.}
  \item \note{Loop corridor reduced from 28 to 23 features; 4 corners serve as natural landmarks.}
\end{itemize}

\logentry{2026-03-05 -- Uniqueness Verification \& User Feedback Summary}
\begin{itemize}
  \item \textbf{User suggestion}: Plain featureless walls will cause SLAM scan-matching to collapse along the corridor axis (classic hallway problem). Added wall-mounted geometric protrusions.
  \item \textbf{User suggestion}: Need a closed-loop corridor for benchmarking (loop closure, continuous laps). Created \texttt{corridor\_loop.sdf}.
  \item \textbf{User suggestion}: Features must not form a repeating pattern --- even periodic features cause SLAM aliasing. Verified and fixed: original features had near-uniform 3.0--3.5m spacing. Redesigned with all-distinct gaps.
  \item \textbf{User question}: AWS RoboMaker Hospital World (\texttt{aws-robotics/aws-robomaker-hospital-world}) --- why not use it?
    \begin{itemize}
      \item Answer: targets Gazebo Classic (7.14+/9.16+), not Ignition Fortress. Different plugin API. ``Humans'' are static meshes, not pedestrians.
      \item Middle ground: use Ignition Fuel models (chairs, carts) for visual fidelity after pipeline is complete. Cosmetic, not scientific.
    \end{itemize}
  \item \textbf{User question}: Why not use \texttt{gazebo\_sfm\_plugin} for pedestrian simulation?
    \begin{itemize}
      \item Answer: designed for Gazebo Classic. Porting to Ignition Fortress requires rewriting plugin API, actor system, and world integration. GT-pose cylinder architecture is sufficient since social stack consumes poses, not LiDAR detections.
    \end{itemize}
  \item \textbf{Scan uniqueness verification (straight corridor)}:
    No translational symmetry exists because no gap value repeats across the full feature set (south gaps: 4.5, 2.9, 5.3, 3.8; north gaps: 2.5, 4.2, 2.3, 3.8, 3.2). Checked mirror positions (x vs $-$x): at x=$-$6 nearest south feature is 2.2m behind vs at x=$+$6 it is 1.5m behind. Cross-wall signatures also differ. Feature size variation (0.15, 0.20, 0.40m widths) adds LiDAR bump-width diversity. No two positions share a scan fingerprint.
  \item \note{L-turn, T-junction, narrow, doorway corridors: corners, width changes, and alcoves serve as powerful unique landmarks. Features there only supplement natural geometry --- aliasing risk is low.}
\end{itemize}

%% ═══════════════════════════════════════════════════════════════════
%%  PHASE 6: SCENARIO SYSTEM + EXPERIMENT RUNNER
%% ═══════════════════════════════════════════════════════════════════

\section*{Phase 6: Scenario System and Experiment Runner}

\subsection*{Design Rationale}

The experiment pipeline must execute $\sim$1,710 simulation trials automatically:
4 controllers $\times$ 6 scenarios $\times$ 30 seeds = 720 main trials, plus
$\sim$990 ablation trials across 7 ablation studies. Each trial involves
launching Gazebo, spawning the robot and pedestrians, running Nav2 with a
specific controller config, collecting metrics, and tearing everything down
cleanly. Manual execution is infeasible.

\textbf{Key architectural decisions:}

\begin{enumerate}
  \item \textbf{Scenario YAML format}: Each of the 6 encounter types is
    defined as a YAML file containing world reference, robot start/goal,
    and pedestrian configurations with randomizable parameters
    (speed, lateral offset, start delay). This separates scenario
    \emph{definition} from \emph{execution}, enabling easy modification
    and extension without touching any code.

  \item \textbf{Seed-based deterministic randomization}: For each trial,
    a seed is passed to \texttt{numpy.random.RandomState(seed)}.
    All pedestrian parameters are sampled from clipped normal distributions:
    $\text{val} = \text{clip}(\mathcal{N}(\mu, \sigma), \text{min}, \text{max})$.
    This ensures:
    \begin{itemize}
      \item Reproducibility: same seed $\Rightarrow$ identical trial
      \item Diversity: different seeds explore the parameter space
      \item Bounds: parameters never exceed configured limits
    \end{itemize}

  \item \textbf{Pedestrian driving via Gazebo set\_pose service}: Since
    Ignition Fortress actors with collision geometry are invisible to
    GPU LiDAR (Phase 0 finding), pedestrians are spawned as dynamic
    \texttt{<model>} entities. A dedicated \texttt{pedestrian\_driver}
    ROS node moves them by calling \texttt{/world/<name>/set\_pose} at
    20Hz, advancing each pedestrian along its waypoint path at the
    configured speed after its start delay elapses.

  \item \textbf{Nav2 controller configs}: 4 standalone YAML files were
    created, each a complete Nav2 parameter set adapted for low-speed
    corridor navigation ($v_{\max}=1.0$ m/s, Ackermann motion model):
    \begin{itemize}
      \item \texttt{nav2\_mppi\_vanilla.yaml}: Baseline MPPI, no social awareness
      \item \texttt{nav2\_mppi\_social\_instant.yaml}: + instantaneous social costmap layer
      \item \texttt{nav2\_mppi\_social\_pred.yaml}: + predictive corridor-frame social layer
      \item \texttt{nav2\_full.yaml}: + predictive layer + CBF safety shield
    \end{itemize}

  \item \textbf{Metrics collection}: The \texttt{metrics\_logger} node
    subscribes to monitoring topics and incrementally accumulates 10 metrics
    per trial. No large buffers or replay bags---each metric is updated
    on message arrival and written to CSV when the trial ends.
\end{enumerate}

\subsection*{6 Encounter Scenarios}

\begin{tabular}{llp{6cm}l}
\hline
\textbf{Scenario} & \textbf{World} & \textbf{Description} & \textbf{Peds} \\
\hline
\texttt{head\_on\_single} & straight & 1 pedestrian walking toward robot & 1 \\
\texttt{head\_on\_group} & straight & 3 pedestrians spanning corridor width & 3 \\
\texttt{doorway\_popout} & doorway & 1 pedestrian emerges from alcove & 1 \\
\texttt{junction\_crossing} & T-junction & 1 pedestrian crosses from branch & 1 \\
\texttt{overtaking} & straight & 1 slow pedestrian ahead, same direction & 1 \\
\texttt{bidirectional} & straight & 4 pedestrians in both directions & 4 \\
\hline
\end{tabular}

\subsection*{Experiment Pipeline Architecture}

\begin{verbatim}
run_all_experiments.sh
  └── run_experiment.py  (per-trial orchestrator)
        ├── scenario_loader.py  (YAML -> TrialConfig)
        ├── sim_experiment.launch.py  (unified ROS 2 launch)
        │     ├── fortress_bringup.launch.py (Gazebo + robot)
        │     ├── EKF state estimation
        │     ├── Nav2 bringup (with controller config)
        │     ├── ros_gz_bridge (scan, imu, clock, poses)
        │     ├── pedestrian_driver (moves spawned models)
        │     ├── gt_people_publisher (Gazebo -> /tracked_people)
        │     ├── gt_collision_detector (GT collision + TTC)
        │     ├── safety_shield_node (if enable_shield=true)
        │     ├── metrics_logger (accumulates trial metrics)
        │     └── goal_sender (sends NavGoal, monitors completion)
        └── CSV output per trial
\end{verbatim}

\subsection*{Metrics Collected Per Trial}

\begin{tabular}{lll}
\hline
\textbf{Metric} & \textbf{Source} & \textbf{Computation} \\
\hline
\texttt{collision\_count} & /collision\_event & Rising-edge count \\
\texttt{min\_separation\_m} & /min\_human\_distance & Global minimum \\
\texttt{ttc\_violations} & /gt\_ttc & Frames with TTC $< 1.0$s \\
\texttt{shield\_interventions\_pct} & /cbf/intervention & \% of True samples \\
\texttt{success} & /trial\_result & Parse ``SUCCESS:...'' \\
\texttt{time\_to\_goal\_s} & /trial\_result & Elapsed time \\
\texttt{path\_length\_m} & /odom & $\sum \sqrt{\Delta x^2 + \Delta y^2}$ \\
\texttt{stuck\_time\_s} & /odom & $\sum dt$ where $|v| < 0.02$ m/s \\
\hline
\end{tabular}

\subsection*{Test Strategy}

Two test suites verify the pipeline's deterministic core:

\textbf{1. Scenario Loader Tests (17 tests):}
\begin{itemize}
  \item \emph{YAML parsing}: All 6 scenario files load without error; fields match expected values.
  \item \emph{Seed reproducibility}: Same seed produces bit-exact identical \texttt{TrialConfig}.
  \item \emph{Seed diversity}: 30 different seeds produce distinct parameter values.
  \item \emph{Bound enforcement}: 100 seeds sampled; all speed/delay/offset values within configured [min, max].
  \item \emph{Waypoint offsetting}: Horizontal walk offsets Y only; vertical walk offsets X only; zero offset preserves waypoints.
  \item \emph{Spawn commands}: Correct model name, coordinates, and SDF path in generated commands.
  \item \emph{Edge cases}: Zero std always returns mean; collinear waypoints; single waypoint; spawn\_z = 0.85.
  \item \emph{Multi-pedestrian diversity}: Group scenario produces different parameters per pedestrian within same seed.
\end{itemize}

\textbf{2. Metrics Computation Tests (31 tests):}
\begin{itemize}
  \item \emph{Collision counting}: Rising-edge detection (sustained collision counts once); rapid toggle counts correctly.
  \item \emph{Min separation}: Tracks global minimum across increasing/decreasing sequences.
  \item \emph{TTC violations}: Threshold at exactly 1.0s (boundary: 1.0s is NOT a violation).
  \item \emph{Shield intervention \%}: Handles 0\%, 50\%, 100\%, and zero-total-samples cases.
  \item \emph{Path integration}: Straight line, diagonal, return trip, stationary, single point.
  \item \emph{Stuck time}: Always moving (0s), always stuck, partial stuck, threshold boundary.
  \item \emph{Result parsing}: SUCCESS, TIMEOUT, ABORTED, REJECTED, TIMEOUT\_NO\_SERVER formats.
  \item \emph{CSV output}: Correct columns, header presence, append mode for multi-trial files.
\end{itemize}

\textbf{All 48 tests pass.}

\subsection*{Bug Fixes During Phase 6}

\begin{itemize}
  \item \textbf{Empty Nav2 config files}: The 4 controller variant YAMLs (\texttt{nav2\_mppi\_vanilla.yaml}, etc.) existed as 0-byte files from Phase 4. They were never actually written. Fully populated with low-speed corridor MPPI parameters.
  \item \textbf{Missing pedestrian driver}: The scenario loader generates spawn commands but nothing was driving the pedestrians along their waypoints. Created \texttt{pedestrian\_driver.py} using Gazebo \texttt{set\_pose} service.
\end{itemize}

\subsection*{Files Created/Modified}

\begin{verbatim}
NEW:
  scenarios/head_on_single.yaml
  scenarios/head_on_group.yaml
  scenarios/doorway_popout.yaml
  scenarios/junction_crossing.yaml
  scenarios/overtaking.yaml
  scenarios/bidirectional.yaml
  corridor_social_nav_bringup/scenario_loader.py
  corridor_social_nav_bringup/pedestrian_driver.py
  corridor_social_nav_bringup/metrics_logger.py  (rewritten from placeholder)
  launch/sim_experiment.launch.py
  scripts/run_experiment.py
  scripts/run_all_experiments.sh
  test/test_scenario_loader.py
  test/test_metrics_computation.py
  rc_model_description/config/nav2_mppi_vanilla.yaml  (populated)
  rc_model_description/config/nav2_mppi_social_instant.yaml  (populated)
  rc_model_description/config/nav2_mppi_social_pred.yaml  (populated)
  rc_model_description/config/nav2_full.yaml  (populated)

MODIFIED:
  corridor_social_nav_bringup/setup.py  (added pedestrian_driver entry point)
  corridor_social_nav_bringup/package.xml  (added ros_gz_interfaces dep)
\end{verbatim}

\section*{Future Work / Nice-to-Haves}
\begin{itemize}
  \item \textbf{AWS RoboMaker Hospital World for visual fidelity}: The \texttt{aws-robomaker-hospital-world} (GitHub: \texttt{aws-robotics/aws-robomaker-hospital-world}) provides a detailed hospital environment with corridors, rooms, furniture, and human models. However:
    \begin{itemize}
      \item It targets Gazebo Classic (7.14+ / 9.16+), NOT Ignition Fortress --- incompatible plugin APIs.
      \item The ``humans'' are static mesh models (ElderLadyPatient, etc.), not moving pedestrians.
      \item Porting the full world is multi-day effort and out of scope.
    \end{itemize}
    \textbf{Middle ground}: Many furniture models (chairs, carts, cabinets) referenced in the hospital world are hosted on Ignition Fuel (\texttt{fuel.ignitionrobotics.org}) and work natively with Fortress. After the experiment pipeline is complete, 2--3 Fuel models can be placed in each corridor as static clutter for more realistic paper screenshots and additional SLAM features. This is cosmetic, not scientific --- the controlled geometric corridors are \emph{better} for a controls paper because confounding variables are minimized.
\end{itemize}

\end{document}
