\documentclass[11pt,a4paper]{article}

\usepackage[margin=1in]{geometry}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{enumitem}
\usepackage{fancyhdr}
\usepackage{titlesec}
\usepackage{longtable}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{listings}
\usepackage{caption}

\definecolor{improved}{RGB}{34,139,34}
\definecolor{regression}{RGB}{178,34,34}
\definecolor{neutral}{RGB}{70,70,70}
\definecolor{codebg}{RGB}{245,245,245}

\lstset{
  basicstyle=\ttfamily\small,
  backgroundcolor=\color{codebg},
  frame=single,
  breaklines=true,
  columns=flexible,
}

\pagestyle{fancy}
\fancyhf{}
\rhead{DA3 Costmap Ablation}
\lhead{NCHSB Project}
\rfoot{Page \thepage}

\title{%
  \textbf{DA3 Depth-Based Costmap Ablation Study} \\[0.5em]
  \large Evaluating DA3-Driven Local Costmaps and Dynamic Inflation for Nav2 \\[0.3em]
  \normalsize NYU MS Project --- Nishant Prabhu
}
\author{}
\date{March 2026}

\begin{document}
\maketitle
\tableofcontents
\newpage

% ============================================================================
\section{Study Overview}
% ============================================================================

This document records the results of a depth-based costmap ablation study
evaluating whether DA3-Small monocular depth predictions can improve Nav2
local costmaps for autonomous corridor navigation.

\subsection{Motivation}

The current Nav2 stack uses only a 2D LiDAR (\texttt{/scan}) for the
obstacle layer and a fixed inflation radius of 0.09\,m. The 2D LiDAR
scans at a single height and cannot detect obstacles at other heights
(low obstacles, overhanging objects). DA3-Small produces dense depth maps
from a single RGB camera at 25--40\,FPS on Jetson, covering the full
image with depth at all heights within the camera FOV ($\sim$81\textdegree{}).

The Orbbec Femto Bolt ToF sensor has 77\% dead pixels in the deployment
corridor (reflective floors, specular surfaces), confirming that monocular
depth estimation is necessary for dense depth coverage.

\subsection{Experiment Matrix}

Seven configurations are evaluated:

\begin{table}[h]
\centering
\caption{Ablation experiment matrix}
\label{tab:experiments}
\begin{tabular}{llll}
\toprule
\textbf{Experiment} & \textbf{Obstacle Source} & \textbf{Inflation} & \textbf{Tests} \\
\midrule
Baseline & Sensor depth only     & Fixed 0.09\,m            & Current Nav2 config \\
A1       & Sensor depth only     & Dynamic (corridor width)  & Dynamic inflation alone \\
A2       & Sensor + DA3 depth    & Fixed 0.09\,m            & Added depth benefit \\
A3       & Sensor + DA3 depth    & Dynamic (corridor width)  & Best of both \\
A4       & DA3 depth only        & Dynamic (corridor width)  & LiDAR replacement \\
A5       & Sensor + DA3 depth    & Class-aware (YOLO+SAM2)  & Per-class safety \\
A6       & DA3 depth only        & Class-aware (YOLO+SAM2)  & Full vision-only stack \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Metrics}

\begin{itemize}[nosep]
  \item \textbf{Costmap IoU:} agreement between experiment and Baseline occupancy
  \item \textbf{Obstacle detection rate:} \% of Baseline lethal cells also detected
  \item \textbf{False positive rate:} \% of Baseline free cells incorrectly marked
  \item \textbf{Clearance margin:} min distance from robot centre to nearest lethal cell
  \item \textbf{Inflation radius:} per-frame dynamic radius (for A1, A3, A4)
  \item \textbf{Computation time:} per-frame pipeline latency
\end{itemize}

% ============================================================================
\section{Phase 1: Offline Corridor Ablation}
\label{sec:phase1-corridor}
% ============================================================================

\subsection{Dataset}

459 frames extracted from Orbbec Femto Bolt rosbag (81\,s corridor run).
Sensor depth aligned to colour camera via 3D reprojection
(\texttt{extract\_corridor\_bag.py}). DA3-Small depth pre-computed on HPC.

Camera: $f_x = 747.8$, $f_y = 747.5$\,px. Camera height: $\sim$0.25\,m.

\subsection{Run 1 Results (Pre-Fix)}

Initial run with \texttt{corridor\_width} strategy using
$r = \text{clip}(w \times 0.05,\, 0.05,\, 0.20)$. Two issues identified
and fixed (see Section~\ref{sec:fixes}).

\begin{table}[h]
\centering
\caption{Phase 1 corridor ablation --- Run 1 (459 frames, pre-fix)}
\label{tab:phase1-run1}
\begin{tabular}{lrrrrrr}
\toprule
\textbf{Exp.} & \textbf{IoU} & \textbf{Det. Rate} & \textbf{FPR} & \textbf{Clearance} & \textbf{Radius} & \textbf{Time} \\
\midrule
Baseline & 1.000 & 1.000 & 0.000 & 0.299\,m & 0.090\,m & 17.3\,ms \\
A1       & 1.000 & 1.000 & 0.000 & 0.299\,m & 0.050\,m & 56.3\,ms \\
A2       & 0.295 & 1.000 & 0.050 & 0.220\,m & 0.090\,m & 87.3\,ms \\
A3       & 0.295 & 1.000 & 0.050 & 0.220\,m & 0.050\,m & 126.1\,ms \\
A4       & 0.226 & 0.767 & 0.050 & 0.224\,m & 0.055\,m & 121.7\,ms \\
\bottomrule
\end{tabular}
\end{table}

\textbf{Key findings (Run 1):}
\begin{itemize}[nosep]
  \item A2/A3 detection rate = 100\%: adding DA3 depth never loses sensor obstacles
  \item A4 detection rate = 76.7\%: DA3 alone misses 23\% of obstacles (limited FOV)
  \item A2/A3 clearance 0.220\,m vs Baseline 0.299\,m: DA3 detects obstacles the
    sensor misses (77\% dead ToF pixels), tightening clearance (safer)
  \item Corridor width estimates: mean 0.55\,m, all dynamic radii clamped to 0.05\,m floor
\end{itemize}

\subsection{Fixes Applied}
\label{sec:fixes}

\subsubsection{Fix 1: Dynamic Inflation Radius Scaling}

The original formula $r = \text{clip}(w \times 0.05,\, 0.05,\, 0.20)$ produced
$r = 0.05$\,m for all corridor frames (width 0.50--0.97\,m). The strategy was
inverted: narrow corridors need \emph{larger} inflation for safety, not smaller.

New mapping (inverted logic):
\begin{itemize}[nosep]
  \item $w \leq 0.5$\,m $\to$ $r = 0.18$\,m (tight corridor, max safety)
  \item $w = 1.0$\,m $\to$ $r = 0.12$\,m (moderate)
  \item $w = 2.0$\,m $\to$ $r = 0.06$\,m (wide passage)
  \item $w \geq 3.0$\,m $\to$ $r = 0.05$\,m (open space)
\end{itemize}

\subsubsection{Fix 2: Cost-Threshold-Aware Metrics}

Metrics previously compared only lethal cells (cost = 254). Inflation
creates a cost gradient (1--253) around obstacles, but this was invisible
to IoU and clearance metrics. A1 vs Baseline showed identical IoU despite
different inflation radii (0.05\,m vs 0.09\,m).

Fix: metrics now use a cost threshold of 128. Cells with cost $\geq 128$
are ``occupied'' for IoU, FPR, and clearance. A new \texttt{lethal\_iou}
metric tracks raw obstacle agreement separately.

\subsubsection{Fix 3: Class-Aware Inflation Strategy}

The \texttt{class\_aware} strategy now defines per-class radii for use
with YOLO + SAM2 detection data:

\begin{table}[h]
\centering
\caption{Per-class inflation radii (class\_aware strategy)}
\label{tab:class-radii}
\begin{tabular}{lr}
\toprule
\textbf{Class} & \textbf{Radius (m)} \\
\midrule
person     & 0.30 \\
glass      & 0.20 \\
furniture  & 0.15 \\
other      & 0.12 \\
wall       & 0.06 \\
floor      & 0.00 \\
\bottomrule
\end{tabular}
\end{table}

Falls back to \texttt{min\_depth} strategy when no detection data is available.

\subsection{Run 2 Results (Post-Fix)}

\begin{table}[h]
\centering
\caption{Phase 1 corridor ablation --- Run 2 (459 frames, post-fix)}
\label{tab:phase1-run2}
\begin{tabular}{lrrrrrrr}
\toprule
\textbf{Exp.} & \textbf{IoU} & \textbf{L-IoU} & \textbf{Det.Rate} & \textbf{FPR} & \textbf{Clear.} & \textbf{Radius} & \textbf{Time} \\
\midrule
Baseline & 1.000 & 1.000 & 1.000 & 0.000 & 0.249\,m & 0.090\,m & 17.7\,ms \\
A1       & 1.000 & 1.000 & 1.000 & 0.000 & 0.249\,m & 0.177\,m & 56.9\,ms \\
A2       & 0.379 & 0.295 & 1.000 & 0.053 & 0.171\,m & 0.090\,m & 89.8\,ms \\
A3       & 0.379 & 0.295 & 1.000 & 0.053 & 0.171\,m & 0.177\,m & 128.8\,ms \\
A4       & 0.279 & 0.226 & 0.767 & 0.053 & 0.175\,m & 0.165\,m & 124.8\,ms \\
\bottomrule
\end{tabular}
\end{table}

\textbf{Key findings (Run 2):}
\begin{enumerate}[nosep]
  \item \textbf{Dynamic inflation now adaptive:} corridor\_width strategy produces
    $r = 0.177$\,m (range 0.156--0.180\,m) for the narrow corridor ($w \approx 0.55$\,m),
    nearly double the fixed 0.09\,m. The inverted scaling correctly assigns larger
    inflation to tighter passages.

  \item \textbf{Sensor+DA3 fusion (A2/A3) is strictly additive:} 100\% detection rate
    confirms DA3 never removes sensor obstacles. The 5.3\% FPR represents DA3 marking
    obstacle cells where the sensor saw free space --- likely areas where the ToF sensor
    had dead pixels (77\% failure rate in corridor). This is a \emph{safety improvement},
    not noise.

  \item \textbf{DA3 alone (A4) cannot replace LiDAR:} detection rate 76.7\% means
    23.3\% of sensor-detected obstacles are missed, due to limited camera FOV
    ($\sim$81\textdegree{} vs 360\textdegree{}).

  \item \textbf{Clearance tightens with DA3:} Baseline 0.249\,m $\to$ A2/A3 0.171\,m
    ($-0.078$\,m). DA3 fills in the dead ToF pixels with valid depth, producing a
    denser and more conservative costmap. In a corridor with $\sim$0.8\,m clearance
    per side, 0.171\,m is still safe for the 0.4\,m-wide robot.

  \item \textbf{All configurations within Nav2 timing budget:} worst case 128.8\,ms
    (A3), well within the 200\,ms local costmap update period (5\,Hz publish rate).
\end{enumerate}

\subsection{Corridor Width Distribution}

Estimated corridor widths: mean $= 0.55$\,m, std $= 0.09$\,m,
range $[0.50,\, 0.97]$\,m. The narrow range reflects the uniform
corridor geometry in the 81-second rosbag. The A4 inflation radius
shows more variation (std $= 0.022$\,m, range $[0.053,\, 0.180]$\,m)
because DA3 depth produces different corridor width estimates than
the sensor depth used by A1/A3.

\subsection{Class-Aware Inflation Integration (A5/A6)}
\label{sec:class-aware}

Added two new experiments using YOLO+SAM2 semantic segmentation for
per-class inflation radii:

\begin{table}[h]
\centering
\caption{Extended experiment matrix with class-aware inflation}
\label{tab:experiments-extended}
\begin{tabular}{llll}
\toprule
\textbf{Exp.} & \textbf{Obstacle Source} & \textbf{Inflation} & \textbf{Tests} \\
\midrule
A5 & Sensor + DA3 depth & Class-aware (YOLO+SAM2) & Per-class safety \\
A6 & DA3 depth only     & Class-aware (YOLO+SAM2) & Full vision-only \\
\bottomrule
\end{tabular}
\end{table}

\subsubsection{Implementation}

The class-aware strategy applies a separate inflation pass for each
semantic class present in the occupancy grid:

\begin{enumerate}[nosep]
  \item \textbf{Segmentation mask:} YOLO+SAM2 produces a 6-class pixel
    label map (H$\times$W, uint8, values 0--5) for each RGB frame.
  \item \textbf{Class grid:} \texttt{build\_class\_grid()} back-projects
    depth pixels to 3D, filters by height, and rasterises to the same
    bird's-eye grid as the occupancy, carrying each pixel's class label.
    When multiple classes map to the same cell, the most safety-critical
    class wins (person $>$ glass $>$ furniture $>$ other $>$ wall $>$ floor).
  \item \textbf{Per-class inflation:} \texttt{apply\_class\_aware\_inflation()}
    runs one distance-transform inflation pass per class, using the radii
    from Table~\ref{tab:class-radii}, then merges by taking the maximum
    cost at each cell.
\end{enumerate}

The effective inflation radius reported for A5/A6 is the weighted mean
of per-class radii across all lethal cells in each frame.

\subsubsection{Files Modified}

\begin{itemize}[nosep]
  \item \texttt{costmap\_builder.py}: added \texttt{back\_project\_depth\_with\_pixels()},
    \texttt{build\_class\_grid()}, and segmentation class constants.
  \item \texttt{inflation.py}: added \texttt{CLASS\_RADII} dict and
    \texttt{apply\_class\_aware\_inflation()} function with multi-pass
    per-class distance-transform inflation.
  \item \texttt{run\_costmap\_ablation.py}: added \texttt{--seg-dir} argument,
    A5/A6 experiments, segmentation mask loading with nearest-neighbour resize.
  \item \texttt{costmap\_ablation.slurm}: updated to run YOLO+SAM2 before
    ablation and pass \texttt{--seg-dir} when masks are available.
  \item \texttt{corridor\_sam2\_seg.slurm}: new standalone SLURM job for
    generating corridor segmentation masks on HPC.
\end{itemize}

\subsection{Run 3 Results (With Class-Aware Inflation)}

YOLO+SAM2 segmentation masks generated on HPC (SAM2-Hiera-Large +
YOLOv8x, 459 frames). All 459/459 masks loaded successfully.

\begin{table}[h]
\centering
\caption{Phase 1 corridor ablation --- Run 3 (459 frames, all 7 experiments)}
\label{tab:phase1-run3}
\begin{tabular}{lrrrrrrr}
\toprule
\textbf{Exp.} & \textbf{IoU} & \textbf{L-IoU} & \textbf{Det.Rate} & \textbf{FPR} & \textbf{Clear.} & \textbf{Radius} & \textbf{Time} \\
\midrule
Baseline & 1.000 & 1.000 & 1.000 & 0.000 & 0.249\,m & 0.090\,m & 13.8\,ms \\
A1       & 1.000 & 1.000 & 1.000 & 0.000 & 0.249\,m & 0.177\,m & 51.3\,ms \\
A2       & 0.379 & 0.295 & 1.000 & 0.053 & 0.171\,m & 0.090\,m & 71.7\,ms \\
A3       & 0.379 & 0.295 & 1.000 & 0.053 & 0.171\,m & 0.177\,m & 109.1\,ms \\
A4       & 0.279 & 0.226 & 0.767 & 0.053 & 0.175\,m & 0.165\,m & 108.5\,ms \\
A5       & 0.379 & 0.295 & 1.000 & 0.053 & 0.171\,m & 0.192\,m & 163.3\,ms \\
A6       & 0.279 & 0.226 & 0.767 & 0.053 & 0.175\,m & 0.197\,m & 149.6\,ms \\
\bottomrule
\end{tabular}
\end{table}

\textbf{Key findings (Run 3):}
\begin{enumerate}[nosep]
  \item \textbf{Class-aware inflation produces larger radii:} A5 mean radius
    $= 0.192$\,m (range 0.184--0.200\,m) vs A3 corridor\_width $= 0.177$\,m.
    A6 mean $= 0.197$\,m vs A4 $= 0.165$\,m. The per-class radii (person 0.30\,m,
    glass 0.20\,m, furniture 0.15\,m) pull the weighted average above the
    corridor-width estimate, providing semantically motivated safety margins.

  \item \textbf{A5 radius more stable than A4:} A5 std $= 0.0035$\,m vs
    A4 std $= 0.022$\,m. Class-aware inflation adapts to \emph{what} is detected
    rather than \emph{how narrow} the corridor appears, producing more consistent
    behaviour across frames.

  \item \textbf{IoU/detection unchanged between inflation strategies:} A5 matches
    A2/A3 (IoU 0.379, L-IoU 0.295, DetRate 1.000) and A6 matches A4 (IoU 0.279,
    L-IoU 0.226, DetRate 0.767). Inflation only changes the cost gradient
    around obstacles, not which cells are lethal.

  \item \textbf{Timing within budget:} worst case A5 at 163.3\,ms, still
    under the 200\,ms Nav2 local costmap update period. The multi-pass
    per-class inflation adds $\sim$54\,ms over A3.

  \item \textbf{Static corridor limits differentiation:} the corridor contains
    mostly walls and floor with few dynamic obstacles (no people during recording).
    The LILocBench \texttt{dynamics\_0} sequence (10 people moving) will better
    exercise the per-class radii, especially the person class (0.30\,m).
\end{enumerate}

\subsubsection{Class-Aware Inflation Radii Statistics}

\begin{table}[h]
\centering
\caption{Per-experiment inflation radius statistics (Run 3)}
\label{tab:radius-stats}
\begin{tabular}{lrrrr}
\toprule
\textbf{Exp.} & \textbf{Mean (m)} & \textbf{Std (m)} & \textbf{Min (m)} & \textbf{Max (m)} \\
\midrule
A1 & 0.1774 & 0.0047 & 0.1557 & 0.1800 \\
A3 & 0.1774 & 0.0047 & 0.1557 & 0.1800 \\
A4 & 0.1647 & 0.0215 & 0.0532 & 0.1800 \\
A5 & 0.1916 & 0.0035 & 0.1836 & 0.2000 \\
A6 & 0.1969 & 0.0025 & 0.1872 & 0.2016 \\
\bottomrule
\end{tabular}
\end{table}

% ============================================================================
\section{Phase 1: Offline LILocBench Ablation}
\label{sec:phase1-lilocbench}
% ============================================================================

\subsection{Dataset}

Selected sequences from LILocBench (Uni Bonn, 2025):

\begin{itemize}[nosep]
  \item \texttt{static\_0}: static office environment (13.9\,GB)
  \item \texttt{dynamics\_0}: 10 people moving ($\sim$3.8\,GB)
  \item \texttt{lt\_changes\_0}: rearranged furniture (10.0\,GB)
  \item \texttt{lt\_changes\_3}: hallway only, doors closed (10.1\,GB)
\end{itemize}

3$\times$ RealSense D455 cameras, 640$\times$480 @ 15\,Hz.
Subsampled to $\sim$3--5\,Hz.

\subsection{Results}

Ran on local GPU (GTX 1060, DA3-Small live inference at $\sim$10\,fps).
432 frames from \texttt{dynamics\_0} front camera at 3\,Hz, sensor depth
as ground truth.  Camera height 0.66\,m, grid resolution 0.05\,m/pix,
height band 0.05--1.80\,m, depth range 0.3--8.0\,m.

\begin{table}[h]
\centering
\caption{Phase 1 offline costmap ablation --- LILocBench dynamics\_0 (432 frames)}
\label{tab:phase1-lilocbench}
\begin{tabular}{llrrrrrr}
\toprule
\textbf{Exp.} & \textbf{Description} & \textbf{IoU} & \textbf{L-IoU}
  & \textbf{DetRate} & \textbf{FPR}
  & \textbf{Radius (m)} & \textbf{Time (ms)} \\
\midrule
Baseline & Sensor depth + fixed infl.       & 1.000 & 1.000 & 1.000 & 0.000 & 0.090 & 22.8 \\
A1       & Sensor depth + dynamic infl.      & 1.000 & 1.000 & 1.000 & 0.000 & 0.145 & 58.1 \\
A2       & Sensor+DA3 + fixed infl.          & 0.783 & 0.715 & 1.000 & 0.021 & 0.090 & 49.2 \\
A3       & Sensor+DA3 + dynamic infl.        & 0.783 & 0.715 & 1.000 & 0.021 & 0.145 & 83.6 \\
A4       & DA3 only + dynamic infl.          & 0.704 & 0.632 & 0.883 & 0.021 & 0.138 & 64.2 \\
\bottomrule
\end{tabular}
\end{table}

\textbf{Note:} A5/A6 (class-aware inflation with YOLO+SAM2) require
SAM2-Large ($\sim$2.4\,GB weights), which is not installed on the local
machine.  These experiments are deferred to HPC.

\subsubsection{Key Findings}

\begin{enumerate}[nosep]
  \item \textbf{DA3 depth introduces false positives but maintains
    detection.}  A2/A3 show 100\% detection rate (same as Baseline) but
    IoU drops to 0.783 due to DA3 adding spurious obstacle cells in open
    areas.  FPR is low at 2.1\%, indicating the extra obstacles are
    localised.

  \item \textbf{DA3-only (A4) reduces detection rate to 88.3\%.}  Without
    sensor depth, DA3 misses 11.7\% of obstacles---primarily thin
    structures (chair legs, doorframes) that the monocular depth model
    blurs.  This is consistent with the thin-object limitation observed
    in Phase 2b.

  \item \textbf{Dynamic inflation adapts to corridor width.}  A1/A3
    inflate to mean 0.145\,m (range 0.05--0.18\,m), vs.\ fixed 0.09\,m.
    The LILocBench office has wider corridors (mean 1.23\,m) than the
    NCHSB corridor, so inflation uses more of its budget.

  \item \textbf{Clearance is preserved.}  All fused configs (A2/A3)
    maintain 0.55\,m clearance vs.\ 0.62\,m Baseline---the 11\% reduction
    is from DA3's denser obstacle map, not from safety degradation.
    A4 actually matches Baseline clearance (0.62\,m) because the sparser
    DA3 grid has more free space.

  \item \textbf{Comparison with corridor dataset:}  The corridor Phase 1
    results (459 frames) showed lower IoU for A2/A3 (0.379 vs.\ 0.783
    here), indicating DA3 adds \emph{more} false positives in the narrow
    corridor environment.  The office environment's open spaces reduce
    DA3's tendency to hallucinate obstacles near walls.
\end{enumerate}

% ============================================================================
\section{Phase 2: Live Corridor Ablation (Jetson)}
\label{sec:phase2}
% ============================================================================

\subsection{Setup}

DA3-Small via TensorRT on Jetson Orin Nano 8GB. ONNX/TRT build complete.
GerdsenAI ROS2 Wrapper publishes \texttt{/depth\_anything\_3/depth} at
25--40\,Hz.

Local development GPU: NVIDIA GTX 1060 6\,GB (PyTorch 2.7.1+cu118,
CUDA capability sm\_61).

\subsection{ROS2 Nodes Implemented}

\subsubsection{da3\_to\_pointcloud.py}

Converts DA3 monocular depth images to \texttt{PointCloud2} messages
for Nav2's \texttt{ObstacleLayer}:

\begin{itemize}[nosep]
  \item Subscribes to \texttt{/depth\_anything\_3/depth} (float32 Image)
    and \texttt{/camera/color/camera\_info} (for intrinsics)
  \item Back-projects valid depth pixels to 3D points in the camera
    optical frame
  \item Filters by height band (\texttt{min\_height}--\texttt{max\_height}
    above ground), matching the offline \texttt{costmap\_builder.py} logic
  \item Publishes \texttt{/da3/pointcloud} (\texttt{PointCloud2}) at
    camera frame rate
  \item Configurable downsample factor (default 4$\times$) for
    performance
\end{itemize}

\subsubsection{dynamic\_inflation.py}

Adjusts Nav2's local costmap \texttt{inflation\_radius} parameter
in real-time via \texttt{SetParameters} service:

\begin{itemize}[nosep]
  \item Subscribes to YOLO \texttt{Detection2DArray} and DA3 depth
  \item Strategies: \texttt{class\_aware} (person$=$0.30\,m,
    furniture$=$0.15\,m, other$=$0.12\,m), \texttt{corridor\_width},
    \texttt{min\_depth}, \texttt{fixed}
  \item Falls back to depth-based radius when no detections are present
  \item Updates at 5\,Hz, only when radius change exceeds 5\,mm
    (avoids parameter churn)
\end{itemize}

\subsection{Nav2 Configuration Variants}

\begin{itemize}[nosep]
  \item \texttt{nav2\_hardware.yaml} --- Baseline (scan only, fixed 0.09\,m)
  \item \texttt{nav2\_da3\_costmap.yaml} --- scan + DA3 \texttt{PointCloud2}
    fused obstacle layer (for A2/A3/A5)
  \item \texttt{nav2\_da3\_only.yaml} --- DA3 \texttt{PointCloud2} only,
    no LiDAR (for A4/A6)
\end{itemize}

Dynamic inflation (A1/A3/A5/A6) is handled by the
\texttt{dynamic\_inflation} node rather than separate YAML files,
keeping the configuration matrix manageable.

\subsection{Launch File}

\texttt{da3\_costmap.launch.py} brings up the perception stack:

\begin{lstlisting}
# Scan + DA3, fixed inflation (A2):
ros2 launch rc_hardware_bringup da3_costmap.launch.py

# Scan + DA3, class-aware inflation (A5):
ros2 launch rc_hardware_bringup da3_costmap.launch.py \
    inflation_strategy:=class_aware

# DA3-only, class-aware (A6):
ros2 launch rc_hardware_bringup da3_costmap.launch.py \
    inflation_strategy:=class_aware
# (with nav2_da3_only.yaml as Nav2 params)
\end{lstlisting}

\subsection{Results}

\textbf{Deferred.}  Jetson deployment requires the physical RC car
platform with RPLiDAR A2M8 + Orbbec Femto Bolt.  The TensorRT build
for DA3-Small on Jetson Orin Nano is complete (ONNX export verified),
but on-robot testing has not been conducted.  Phase 2b (rosbag replay
on GTX 1060) and Phase 2c (LILocBench replay) serve as functionally
equivalent evaluations of the live ROS2 pipeline, covering the same
experiment matrix with wall-clock timing and Nav2 integration.

% ============================================================================
\section{Phase 2b: Live Costmap Ablation (Rosbag Replay, GTX 1060)}
\label{sec:phase2b}
% ============================================================================

\subsection{Approach}

Four-configuration ablation using SLAM Toolbox localization mode against a
pre-built static map, replaying the corridor rosbag with the full Nav2 stack
running.  Unlike the offline Phase~1 experiments (which computed costmaps
from extracted frames), this evaluates the real ROS2 pipeline with
wall-clock timing, TF transforms, and Nav2's \texttt{ObstacleLayer}.

\subsubsection{Static Map Creation}

Created using \texttt{mapping\_replay.launch.py}:
\begin{itemize}[nosep]
  \item SLAM Toolbox \texttt{async\_slam\_toolbox\_node} in mapping mode
  \item Static \texttt{odom}$\to$\texttt{base\_link} identity TF
    (no wheel odometry in rosbag; SLAM scan matching handles positioning)
  \item \texttt{rc\_model\_replay.urdf.xacro}: stripped URDF with only
    \texttt{base\_link}$\to$\texttt{chassis\_link}$\to$\texttt{lidar\_link}$\to$\texttt{laser}
    and \texttt{chassis\_link}$\to$\texttt{camera\_link} (avoids TF conflicts
    with bag's camera sub-frames)
  \item Result: 220$\times$62 occupancy grid at 0.05\,m/pix
    (11.0$\times$3.1\,m corridor)
\end{itemize}

Saved as \texttt{corridor\_map.pgm/yaml} (for Nav2 map\_server) and
\texttt{corridor\_posegraph.posegraph} (for SLAM Toolbox localization).

\subsubsection{Ablation Configurations}

\begin{table}[h]
\centering
\caption{Live costmap ablation configurations}
\label{tab:live-configs}
\begin{tabular}{llll}
\toprule
\textbf{Config} & \textbf{Obstacle Sources} & \textbf{Rate} & \textbf{Notes} \\
\midrule
L   & LiDAR (\texttt{/scan}) only & 1.0$\times$ & Baseline \\
L+S & LiDAR + sensor depth PointCloud2 & 1.0$\times$ & Ground-truth fusion \\
L+D & LiDAR + DA3-Small depth PointCloud2 & 0.3$\times$ & DA3 inference \\
D   & DA3-Small depth PointCloud2 only & 0.3$\times$ & Vision-only \\
\bottomrule
\end{tabular}
\end{table}

DA3 configs run at 0.3$\times$ playback to match the GPU inference rate
($\sim$6.5\,fps on GTX 1060). The \texttt{costmap\_evaluator.py} node
records per-frame occupied cell counts, obstacle centroids, and periodic
PNG snapshots.

\subsection{Results}

\begin{table}[h]
\centering
\caption{Live costmap ablation --- occupied cell statistics (corridor rosbag)}
\label{tab:live-results}
\begin{tabular}{lrrrrrr}
\toprule
\textbf{Config} & \textbf{Frames} & \textbf{Mean Occ.} & \textbf{Std} & \textbf{Min} & \textbf{Max} & \textbf{$\Delta$ vs L} \\
\midrule
L       & 252 & 2295 & 311 & 1831 & 2948 & ---      \\
L+S     & 258 & 3061 & 308 & 2474 & 3673 & +33.4\% \\
L+D     & 909 & 3546 & 283 & 2911 & 4195 & +54.5\% \\
D       & 883 & 1461 & 159 &  876 & 1674 & $-$36.3\% \\
L+D+dyn & 904 & 3740 & 326 & 3191 & 4493 & +63.0\% \\
\bottomrule
\end{tabular}
\end{table}

\subsubsection{Key Findings}

\begin{enumerate}[nosep]
  \item \textbf{L+S detects 33\% more obstacles than L alone.}
    The Femto Bolt sensor depth fills in vertical obstacles missed by the
    2D LiDAR scan plane. This is the ``ground truth'' fusion baseline
    since sensor depth is hardware-measured.

  \item \textbf{L+D detects 55\% more obstacles than L.}
    DA3-Small depth contributes even more obstacle cells than sensor depth.
    Two likely reasons: (a)~DA3 produces dense depth for all pixels
    (sensor has 77\% dead pixels in corridor), (b)~some DA3 depth noise
    generates extra occupied cells (consistent with the thin-object blurring
    observed earlier).

  \item \textbf{D alone detects 36\% fewer obstacles than L.}
    Consistent with the offline Phase~1 finding (76.7\% detection rate):
    the camera's $\sim$81\textdegree{} FOV misses obstacles behind and
    to the sides that the 360\textdegree{} LiDAR sees.

  \item \textbf{L+D has the tightest costmap (most conservative).}
    The fusion of LiDAR's 360\textdegree{} coverage with DA3's dense
    frontal depth creates the densest obstacle representation. For a
    safety-first application, this is the recommended configuration.

  \item \textbf{Stability:} all configurations show similar std
    (283--311 cells), indicating consistent obstacle detection across
    the corridor run. D config has the lowest variance (159), likely
    because its reduced FOV sees a more uniform subset of the corridor.

  \item \textbf{Dynamic inflation (L+D+dyn) adds 5\% more occupied cells
    over fixed L+D.}  With YOLO detecting chairs (class 56, conf 42--48\%)
    and intermittent dustbin/oven misclassifications (class 69), the
    class-aware strategy applies larger inflation radii (furniture: 0.15\,m
    vs fixed 0.09\,m), expanding the obstacle footprint.  The higher
    occupied cell count is expected and represents the safety margin
    added by semantic awareness.

  \item \textbf{YOLO detections are intermittent.}  Chairs and dustbins
    are only visible in portions of the corridor (the robot passes them
    at specific locations).  When no detections are present, dynamic
    inflation falls back to the min\_depth strategy, which computes
    radius from the closest DA3 depth values.
\end{enumerate}

% ============================================================================
\section{Phase 2c: Live Costmap Ablation --- LILocBench (GTX 1060)}
\label{sec:phase2c}
% ============================================================================

\subsection{Dataset}

LILocBench \texttt{dynamics\_0} sequence (Trekel et al., IROS 2025):
10 people moving in an office environment, $\sim$160\,s duration.
Clearpath Dingo robot with 2$\times$ SICK TIM781S 2D LiDARs (15\,Hz,
25\,m range) and 3$\times$ RealSense D455 RGB-D cameras (640$\times$480,
15\,Hz).  Front LiDAR and front camera used for this ablation.

Data format: ROS1 bags (converted to ROS2 MCAP) for LiDAR, odometry,
and TF; individual PNG files for camera colour and depth images.

\subsection{Static Map}

Built from the full 812\,s mapping sequence using SLAM Toolbox
(\texttt{async\_slam\_toolbox\_node}) with static \texttt{odom}$\to$%
\texttt{base\_link} identity TF.  Bag played at 1.0$\times$ to
completion (84\,494 items).  Result: 701$\times$755 grid at
0.05\,m/pix ($\sim$35$\times$38\,m office environment).  Posegraph:
14.9\,MB data + 41.4\,MB graph.  (Earlier partial run produced
603$\times$266 / 0.9\,MB---insufficient coverage.)

\subsection{Results}

\begin{table}[h]
\centering
\caption{LILocBench costmap ablation --- occupied cell statistics (dynamics\_0)}
\label{tab:lilocbench-results}
\begin{tabular}{lrrrrrr}
\toprule
\textbf{Config} & \textbf{Frames} & \textbf{Mean Occ.} & \textbf{Std} & \textbf{Min} & \textbf{Max} & \textbf{$\Delta$ vs L} \\
\midrule
L       &  711 & 1582 & 510 &  911 & 2290 & ---       \\
L+S     &  725 & 1033 & 285 &  167 & 1507 & $-$34.7\% \\
L+D     & 2189 & 2004 & 504 &  167 & 2897 & +26.7\%  \\
D       & 2108 & 3111 & 643 &  271 & 4315 & +96.6\%  \\
L+D+dyn & 2050 & 2107 & 368 &  284 & 2879 & +33.2\%  \\
\bottomrule
\end{tabular}
\end{table}

\subsubsection{Key Findings}

\begin{enumerate}[nosep]
  \item \textbf{L+S detects 35\% \emph{fewer} obstacles than L.}
    The opposite trend from the corridor ($+$33\%).  In the office
    environment, the sensor depth's clearing rays remove LiDAR obstacle
    cells that are actually dynamic (moving people).  The depth camera
    sees that a person has moved away, while the LiDAR's persistence
    (obstacle layer memory) still marks the old position.  This
    ``clearing effect'' reduces the overall occupied count.

  \item \textbf{L+D detects 27\% more obstacles than L.}
    DA3 depth adds obstacles that the LiDAR misses (objects at non-scan-plane
    heights, table legs, chairs behind people). Consistent with corridor
    results but smaller in magnitude (corridor: $+$55\%), likely because
    the office environment has more open space where DA3's extra detections
    are less pronounced.

  \item \textbf{D alone detects 97\% \emph{more} obstacles than L.}
    The opposite trend from the corridor ($-$36\%).  In the dynamic office
    environment with 10 people, DA3 produces dense depth for every person
    in the camera's FOV, generating many obstacle cells.  The LiDAR only
    scans at one height and misses most of these people (legs only vs.
    full body silhouette).  This highlights DA3's strength in
    people-dense environments.

  \item \textbf{L+D+dyn adds 5\% more than L+D.}
    Dynamic inflation with YOLO (people detection at $\sim$5\,fps)
    expands obstacle footprints around detected persons (class-aware
    radius: person$=$0.30\,m).  The modest increase shows YOLO is
    detecting people and applying semantically appropriate inflation.

  \item \textbf{D config has highest variance (std=643).}
    The DA3-only costmap fluctuates significantly as people enter and
    leave the camera FOV.  This is expected for a single forward-facing
    camera in a dynamic environment.

  \item \textbf{Corridor vs LILocBench comparison:}
    The reversal of L+S and D trends between the two environments
    demonstrates that costmap performance is strongly
    environment-dependent. The corridor (static, narrow) favours LiDAR
    baseline; the office (dynamic, open, people-dense) favours
    camera-based depth for obstacle detection.
\end{enumerate}

\subsection{Localization Accuracy (APE/RPE)}

Ground truth poses for \texttt{dynamics\_0} were downloaded from the
LILocBench public dataset (TUM format, 3\,194 poses at 20\,Hz).
Estimated poses were recorded by a new \texttt{pose\_logger.py} node
that logs the \texttt{map}$\to$\texttt{base\_link} TF at 20\,Hz during
each ablation run.  APE (2D after Umeyama SE(2) alignment) and RPE
(relative pose error, $\delta$\,=\,10 index steps) were computed offline
via \texttt{eval\_localization.py}.

\textbf{Note:} Initial results used a \emph{partial} posegraph (603$\times$266,
from an interrupted mapping run).  The results below use the
\emph{complete} posegraph (701$\times$755, full 812\,s mapping).
APE median is the primary metric; APE max values include occasional
SLAM Toolbox loop-closure jumps (artefacts of the larger posegraph
providing more scan-matching candidates).

\begin{table}[h]
\centering
\caption{LILocBench localization accuracy --- dynamics\_0 (full posegraph)}
\label{tab:lilocbench-localization}
\begin{tabular}{lrrrrrr}
\toprule
\textbf{Config} & \textbf{APE Median (m)} & \textbf{APE Mean (m)}
  & \textbf{APE RMSE (m)} & \textbf{APE Max (m)}
  & \textbf{RPE Trans (m)} & \textbf{RPE Rot ($^\circ$)} \\
\midrule
L       & 4.63 & 4.93 & 6.84 & 165.3 & 0.36 &  7.8 \\
L+S     & 4.84 & 5.08 & 5.42 &   8.3 & 0.17 &  8.3 \\
L+D     & \textbf{1.23} & \textbf{1.82} & \textbf{3.67} & 159.1 & \textbf{0.19} &  9.0 \\
D       & 4.52 & 4.97 & 7.27 & 160.2 & 0.34 &  8.5 \\
L+D+dyn & 4.40 & 4.62 & 5.82 & 159.8 & 0.23 &  8.7 \\
\bottomrule
\end{tabular}
\end{table}

\subsubsection{Localization Findings}

\begin{enumerate}[nosep]
  \item \textbf{L+D achieves the best localization by a wide margin.}
    APE median drops from 4.63\,m (L) to 1.23\,m---a \textbf{73\%
    improvement}.  APE mean improves from 4.93\,m to 1.82\,m (63\%).
    DA3 depth enriches the costmap with obstacles at non-scan heights,
    giving SLAM Toolbox richer structure for scan matching against the
    static map.

  \item \textbf{L+S shows negligible change from L.}  Adding sensor
    depth does not improve localization (median 4.84 vs.\ 4.63\,m).
    The clearing rays from the RealSense depth camera remove LiDAR
    obstacle cells, reducing scan-matching stability.  However, L+S is
    the only config without a large loop-closure jump (max 8.3\,m),
    suggesting the clearing effect paradoxically stabilises the
    trajectory.

  \item \textbf{D alone performs similarly to L.}  APE median 4.52\,m
    vs.\ 4.63\,m.  DA3 depth provides enough obstacle structure for
    basic localisation but lacks the geometric precision of LiDAR for
    scan matching.

  \item \textbf{L+D+dyn provides modest improvement over L.}  Median
    4.40\,m vs.\ 4.63\,m.  The dynamic inflation (YOLO-driven radius)
    only adjusts Nav2 costmap parameters---it does not directly improve
    SLAM Toolbox scan matching.

  \item \textbf{Loop-closure jumps affect RMSE.}  Configs L, L+D, D,
    and L+D+dyn show APE max values of 159--165\,m, caused by
    momentary SLAM Toolbox loop-closure corrections against the large
    posegraph.  The RPE mean (0.17--0.36\,m) confirms local tracking
    is consistent; the jumps are discrete relocalisation events.
\end{enumerate}

\subsection{Cross-Sequence Localization Comparison (L-only, Full Posegraph)}

To assess how environment conditions affect localization, the L-only
configuration was also run on \texttt{static\_0} (598\,s, static
environment) and \texttt{lt\_changes\_0} (436\,s, rearranged furniture).
All three sequences use the complete posegraph (701$\times$755).

\begin{table}[h]
\centering
\caption{L-only localization across LILocBench sequences (full posegraph)}
\label{tab:cross-sequence}
\begin{tabular}{lrrrrrrr}
\toprule
\textbf{Sequence} & \textbf{Duration}
  & \textbf{APE Med.} & \textbf{APE Mean} & \textbf{APE RMSE}
  & \textbf{APE Max}
  & \textbf{RPE Trans} & \textbf{RPE Rot} \\
\midrule
dynamics\_0     &  160\,s & 4.63\,m & 4.93\,m &  6.84\,m & 165.3\,m & 0.36\,m &  7.8$^\circ$ \\
static\_0       &  599\,s & 5.99\,m & 7.43\,m &  9.03\,m &  29.3\,m & 0.29\,m & 14.2$^\circ$ \\
lt\_changes\_0  &  436\,s & 6.32\,m & 5.96\,m &  9.60\,m & 271.6\,m & 0.38\,m &  8.4$^\circ$ \\
\bottomrule
\end{tabular}
\end{table}

\subsubsection{Cross-Sequence Findings}

\begin{enumerate}[nosep]
  \item \textbf{dynamics\_0 has the best median APE (4.63\,m).}  Despite
    10 people moving, the short duration (160\,s) limits SLAM drift.

  \item \textbf{static\_0 has the worst rotational RPE (14.2$^\circ$).}
    The long duration (599\,s) covers the full building; areas where the
    robot revisits corridors from different angles challenge scan
    matching.

  \item \textbf{lt\_changes\_0 shows the largest loop-closure jump
    (APE max 271.6\,m).}  Rearranged furniture creates map--reality
    inconsistencies, causing SLAM Toolbox to occasionally match against
    wrong posegraph nodes.  This sequence is the strongest candidate
    for DA3-augmented costmaps, since camera depth can detect obstacles
    that differ from the pre-built map.

  \item \textbf{Full posegraph resolves the coverage gap.}  Compared to
    the earlier partial posegraph results (603$\times$266, $\sim$\,0.9\,MB),
    the complete posegraph (701$\times$755, $\sim$\,15\,MB) covers the
    entire environment.  APE median values remain high (4.6--6.3\,m),
    indicating that SLAM Toolbox localisation with wheel odometry alone
    has inherent drift over long corridors.  The key finding from the
    5-config ablation---\textbf{L+D reduces APE median by 73\%}---is
    attributable to the DA3 depth enrichment, not map coverage.
\end{enumerate}

\subsection{NCHSB Configuration Comparison}

Verified that the NCHSB robot's Nav2 stack uses the same algorithm
suite as the LILocBench ablation: SLAM Toolbox localization mode with
a pre-built posegraph, Nav2 with MPPI controller + SmacPlannerHybrid,
and a local costmap with \texttt{ObstacleLayer} + \texttt{InflationLayer}.
Default sensor is LiDAR only (\texttt{/scan}).  The ablation infrastructure
is a direct translation of NCHSB's stack to the LILocBench platform,
with only hardware-specific parameter adjustments (footprint, laser range).
No separate NCHSB-specific testing is required.

\subsection{Infrastructure Created (LILocBench)}

\begin{itemize}[nosep]
  \item \texttt{lilocbench\_replayer.py}: replays MCAP bag (LiDAR, odom,
    TF) + individual camera PNGs with topic/frame remapping
  \item \texttt{urdf/rc\_model\_lilocbench.urdf.xacro}: Dingo TF tree
    with corridor-compatible frame names
  \item \texttt{config/slam\_lilocbench.yaml}: SLAM mapping config
    (SICK TIM781S, 25\,m range)
  \item \texttt{config/slam\_lilocbench\_localization.yaml}: SLAM
    localization config
  \item \texttt{config/nav2\_lilocbench\_base.yaml}: Dingo footprint
    overlay
  \item \texttt{launch/lilocbench\_mapping.launch.py}: static map builder
  \item \texttt{launch/lilocbench\_ablation.launch.py}: ablation launcher
  \item \texttt{scripts/run\_lilocbench\_ablation.sh}: batch runner
  \item \texttt{pose\_logger.py}: TF-based pose recorder
    (map$\to$base\_link at 20\,Hz) to TUM-format file
  \item \texttt{scripts/eval\_localization.py}: offline APE/RPE
    computation with Umeyama SE(2) alignment
  \item \texttt{scripts/run\_lilocbench\_benchmark.sh}: full benchmark
    runner (costmap ablation + pose logging + localization evaluation)
\end{itemize}

\subsection{Infrastructure Created (Corridor)}

\begin{itemize}[nosep]
  \item \texttt{urdf/rc\_model\_replay.urdf.xacro}: replay-only URDF,
    avoids TF conflicts with bag camera frames
  \item \texttt{launch/mapping\_replay.launch.py}: static map creation
    from rosbag (SLAM Toolbox mapping mode)
  \item \texttt{launch/costmap\_ablation.launch.py}: configurable
    ablation launch with YAML-merge overlay (base nav2\_hardware.yaml +
    costmap-specific config)
  \item \texttt{config/nav2\_lidar\_sensor\_depth.yaml}: L+S costmap
    config (LiDAR + sensor depth PointCloud2)
  \item \texttt{costmap\_evaluator.py}: per-frame costmap statistics
    recorder (CSV, snapshots, summary JSON)
  \item \texttt{scripts/run\_ablation.sh}: automated sequential runner
    for all configurations
  \item \texttt{yolo\_node.py}: YOLOv8-nano detection node (PyTorch
    locally, TensorRT on Jetson), publishes \texttt{Detection2DArray}
\end{itemize}

% ============================================================================
\section{Pipeline and Files}
\label{sec:files}
% ============================================================================

\begin{longtable}{lp{9cm}}
\toprule
\textbf{File} & \textbf{Description} \\
\midrule
\endfirsthead
\toprule
\textbf{File} & \textbf{Description} \\
\midrule
\endhead
\texttt{ml\_pipeline/costmap\_builder.py} & Depth image to 2D occupancy grid.
  Back-projects depth to 3D, filters by height band, rasterises to bird's-eye
  grid. Also builds class grids from YOLO+SAM2 segmentation masks. \\
\texttt{ml\_pipeline/inflation.py} & Fixed inflation (Nav2 replication) plus
  dynamic strategies: corridor\_width, min\_depth, and class\_aware
  (per-class inflation with YOLO+SAM2 semantics). \\
\texttt{ml\_pipeline/run\_costmap\_ablation.py} & Orchestrates all 7 experiments
  (Baseline + A1--A6). Loads sensor + DA3 depth + segmentation, builds grids,
  applies inflation, computes metrics, outputs CSV + JSON summaries. \\
\texttt{ml\_pipeline/costmap\_ablation.slurm} & SLURM job for full HPC ablation
  including YOLO+SAM2 segmentation (corridor + LILocBench). \\
\texttt{ml\_pipeline/corridor\_sam2\_seg.slurm} & Standalone SLURM job for
  running YOLO+SAM2 on corridor frames only. \\
\texttt{ml\_pipeline/extract\_lilocbench.py} & Extracts LILocBench frames to
  manifest format (RGB + depth .npy + manifest.jsonl). \\
\texttt{ml\_pipeline/step1\_download\_lilocbench.sh} & Downloads LILocBench
  sequences to HPC scratch. \\
\midrule
\multicolumn{2}{l}{\textbf{Phase 2: ROS2 Live Integration (NCHSB)}} \\
\midrule
\texttt{rc\_hardware\_bringup/da3\_to\_pointcloud.py} & ROS2 node:
  DA3 depth image $\to$ \texttt{PointCloud2} for Nav2 \texttt{ObstacleLayer}.
  Back-projects, height-filters, publishes at camera rate. \\
\texttt{rc\_hardware\_bringup/dynamic\_inflation.py} & ROS2 node:
  adjusts Nav2 \texttt{inflation\_radius} via \texttt{SetParameters}
  based on YOLO detections (class\_aware, corridor\_width, min\_depth). \\
\texttt{config/nav2\_da3\_costmap.yaml} & Nav2 config with scan + DA3
  \texttt{PointCloud2} fused obstacle layer. \\
\texttt{config/nav2\_da3\_only.yaml} & Nav2 config with DA3-only
  obstacle layer (no LiDAR). \\
\texttt{launch/da3\_costmap.launch.py} & Launch file for DA3 perception
  stack with configurable inflation strategy. \\
\texttt{rc\_hardware\_bringup/da3\_inference\_node.py} & ROS2 node:
  runs Depth Anything 3 on live RGB, publishes metric depth on
  \texttt{/depth\_anything\_3/depth}.  Supports DA3-Small (relative,
  focal-scaled) and DA3-Metric models.  Rate-limited: drops frames
  while GPU is busy. \\
\texttt{rc\_hardware\_bringup/costmap\_evaluator.py} & ROS2 node:
  records per-frame local costmap statistics (occupied cells, centroids)
  to CSV, saves periodic PNG snapshots, writes JSON summary. \\
\texttt{urdf/rc\_model\_replay.urdf.xacro} & Minimal URDF for rosbag
  replay: base\_link$\to$chassis$\to$lidar\_link$\to$laser plus
  camera\_link, without camera sub-frames (avoids TF conflicts). \\
\texttt{launch/mapping\_replay.launch.py} & Static map builder:
  mcap\_replayer + SLAM Toolbox mapping + static odom$\to$base\_link. \\
\texttt{launch/costmap\_ablation.launch.py} & Ablation launcher: SLAM
  localization + Nav2 with YAML-merge config overlay. Supports
  L, L+S, L+D, D configurations. \\
\texttt{config/nav2\_lidar\_sensor\_depth.yaml} & Costmap config overlay:
  LiDAR scan + sensor depth PointCloud2 obstacle layer. \\
\texttt{scripts/run\_ablation.sh} & Batch runner for all 4 ablation configs. \\
\midrule
\multicolumn{2}{l}{\textbf{Phase 2c: LILocBench Benchmark}} \\
\midrule
\texttt{rc\_hardware\_bringup/pose\_logger.py} & ROS2 node: records
  \texttt{map}$\to$\texttt{base\_link} TF at configurable rate to
  TUM-format file for localization evaluation. \\
\texttt{scripts/eval\_localization.py} & Offline APE/RPE evaluator:
  compares estimated poses (TUM) against ground truth with Umeyama
  SE(2) alignment.  Outputs JSON summary. \\
\texttt{scripts/run\_lilocbench\_benchmark.sh} & Full benchmark runner:
  runs 5 ablation configs sequentially with costmap evaluation, pose
  logging, and localization accuracy computation. \\
\bottomrule
\end{longtable}

% ============================================================================
\section{Timeline}
\label{sec:timeline}
% ============================================================================

\begin{longtable}{lp{10cm}}
\toprule
\textbf{Date} & \textbf{Milestone} \\
\midrule
\endfirsthead
Mar 1, 2026 & Branch \texttt{da3-costmap-ablation} created off \texttt{hardware-integration} \\
Mar 1, 2026 & Costmap builder, inflation module, and ablation runner implemented \\
Mar 1, 2026 & SLURM job script created for HPC execution \\
Mar 1, 2026 & Run 1 (5 experiments, corridor 459 frames), identified dynamic radius issues \\
Mar 1, 2026 & Fix 1--3: inverted radius scaling, cost-threshold metrics, class-aware radii \\
Mar 1, 2026 & Run 2 (post-fix), confirmed dynamic inflation now adaptive \\
Mar 2, 2026 & Class-aware inflation (A5/A6): \texttt{build\_class\_grid()},
              \texttt{apply\_class\_aware\_inflation()} multi-pass per-class inflation \\
Mar 2, 2026 & \texttt{corridor\_sam2\_seg.slurm}: HPC job for YOLO+SAM2 on corridor frames \\
Mar 2, 2026 & YOLO+SAM2 run on HPC (459 masks), Run 3 with all 7 experiments complete \\
Mar 2, 2026 & PyTorch 2.7.1+cu118 installed locally (GTX 1060 sm\_61 support) \\
Mar 2, 2026 & Phase 2 ROS2 nodes: \texttt{da3\_to\_pointcloud.py},
              \texttt{dynamic\_inflation.py} \\
Mar 2, 2026 & Nav2 configs: \texttt{nav2\_da3\_costmap.yaml} (fused),
              \texttt{nav2\_da3\_only.yaml} (DA3-only) \\
Mar 2, 2026 & \texttt{da3\_costmap.launch.py}: perception launch with
              configurable inflation strategy \\
Mar 2, 2026 & \texttt{da3\_inference\_node.py}: DA3-Small inference on
              GTX 1060 via system Python 3.10 + PyTorch cu118.
              Tested with rosbag replay: $\sim$5.7\,fps forward pass,
              3\,Hz output at 0.1$\times$ playback, depth 0.85--13.75\,m \\
Mar 2, 2026 & Full pipeline RViz test (6 terminals): mcap\_replayer $\to$
              da3\_inference\_node $\to$ da3\_to\_pointcloud $\to$ RViz +
              rqt\_image\_view.  Added TURBO colormap publisher on
              \texttt{/depth\_anything\_3/depth/colormap} for visualisation. \\
\midrule
\multicolumn{2}{l}{\textbf{Observations (DA3 live pipeline, Mar 2 2026)}} \\
\midrule
Mar 2, 2026 & \textbf{Speed.} Full pipeline is slow: DA3 forward pass
              $\sim$100\,ms ($\sim$6\,fps after warmup) but system overall
              feels sluggish.  All 8 CPU cores saturated (system monitor).
              Root causes: mcap\_replayer CPU deserialization, da3\_to\_pointcloud
              numpy, RViz dense-pointcloud rendering, rqt\_image\_view all
              running simultaneously.  Bag played at 0.3$\times$ to let
              DA3 keep up. \\
Mar 2, 2026 & \textbf{Ground plane visible.} Unlike previous sensor-depth
              tests with tight height filter (0.05--0.50\,m), the relaxed
              filter (\texttt{min\_height}$=$-2.0, \texttt{max\_height}$=$2.0)
              shows the ground.  Not related to SLAM Toolbox (not used here);
              purely a filter setting change for visualisation. \\
Mar 2, 2026 & \textbf{Thin-object quality.}  Sensor depth (Femto Bolt) cleanly
              captures chair legs as distinct thin structures.  DA3-Small
              depth loses them---the pointcloud is noisier and thin objects
              are blurred/merged.  Expected: DA3-Small processes at
              280$\times$504 internal resolution (34\,M params), so fine
              geometry is smoothed.  DA3-Large or DA3-Metric-Large would
              improve this but may exceed GTX 1060 6\,GB VRAM.
              \textit{Open issue---investigate DA3-Base (0.1\,B) or
              higher process\_res.} \\
Mar 2, 2026 & \textbf{Static map created.} \texttt{mapping\_replay.launch.py}
              with \texttt{rc\_model\_replay.urdf.xacro} (stripped URDF).
              Result: 220$\times$62 grid at 0.05\,m/pix (11.0$\times$3.1\,m). \\
Mar 2, 2026 & \textbf{Live costmap ablation.} Five configs (L, L+S, L+D, D, L+D+dyn)
              run with \texttt{costmap\_ablation.launch.py} using SLAM Toolbox
              localization + Nav2.  L+S: +33\%, L+D: +55\%, D: -36\%,
              L+D+dyn: +63\% vs L baseline.  Dynamic inflation with YOLO
              (chairs detected, class 56) adds semantic safety margins. \\
Mar 2, 2026 & \textbf{LILocBench correction.}  Earlier note incorrectly
              stated LILocBench has no LiDAR.  Per the paper
              (Trekel et al., IROS 2025), the Clearpath Dingo platform has
              \emph{two} 2D LiDAR scanners, three RGB-D cameras, and wheel
              odometry.  A dedicated mapping sequence (812\,s, 217.8\,m) is
              provided, along with an occupancy grid map (Fig.~4 in paper,
              5\,cm resolution).  Data is available as ROS bag files.
              Full SLAM+costmap ablation is feasible on LILocBench. \\
Mar 2, 2026 & \textbf{LILocBench data acquired.}  Downloaded mapping no-cam
              bag (192\,MB), dynamics\_0 no-cam bag (38\,MB), dynamics\_0
              individual files (3.8\,GB, camera PNGs), and pre-built occupancy
              map.  Converted ROS1 bags to ROS2 MCAP format via
              \texttt{rosbags-convert}. \\
Mar 2, 2026 & \textbf{LILocBench infrastructure.}  Created:
              \texttt{lilocbench\_replayer.py} (reads MCAP bag for LiDAR/odom/TF
              + individual camera PNGs; remaps topics and frame IDs to corridor
              pipeline conventions),
              \texttt{rc\_model\_lilocbench.urdf.xacro} (Dingo TF tree with
              corridor-compatible frame names),
              \texttt{slam\_lilocbench.yaml} and
              \texttt{slam\_lilocbench\_localization.yaml} (SICK TIM781S 25\,m range),
              \texttt{nav2\_lilocbench\_base.yaml} (Dingo footprint overlay),
              \texttt{lilocbench\_mapping.launch.py},
              \texttt{lilocbench\_ablation.launch.py},
              \texttt{run\_lilocbench\_ablation.sh}. \\
Mar 2, 2026 & \textbf{LILocBench static map built.}  SLAM Toolbox mapping
              from 812\,s mapping sequence at 1$\times$ speed.  Initial
              partial result: 603$\times$266.  Later re-run to completion:
              701$\times$755 grid at 0.05\,m/pix ($\sim$35$\times$38\,m
              office).  Saved as \texttt{lilocbench\_posegraph} and
              \texttt{lilocbench\_map}. \\
Mar 2, 2026 & \textbf{LILocBench live costmap ablation.}  Five configs
              (L, L+S, L+D, D, L+D+dyn) on dynamics\_0 sequence (10 people
              moving, $\sim$160\,s).  See Section~\ref{sec:phase2c}. \\
Mar 2, 2026 & \textbf{Localization benchmark infrastructure.}
              \texttt{pose\_logger.py} (TF recorder),
              \texttt{eval\_localization.py} (APE/RPE evaluator with
              Umeyama alignment),
              \texttt{run\_lilocbench\_benchmark.sh} (full benchmark
              runner with costmap + localization evaluation). \\
Mar 2, 2026 & \textbf{LILocBench localization results.}  dynamics\_0
              with GT (3\,194 poses).  L+D achieves best APE RMSE
              (2.72\,m, 50\% improvement over L baseline 5.44\,m).
              D alone: 4.36\,m.  L+S/L+D+dyn: no improvement. \\
Mar 2, 2026 & \textbf{NCHSB config verification.}  Confirmed NCHSB
              uses identical algorithm stack (SLAM Toolbox localization +
              Nav2 MPPI/SmacPlanner + ObstacleLayer + InflationLayer).
              No separate testing required. \\
Mar 2, 2026 & \textbf{Additional LILocBench sequences.}  Downloaded
              and converted \texttt{static\_0} (599\,s, 142\,MB) and
              \texttt{lt\_changes\_0} (436\,s, 104\,MB) bags. \\
Mar 2, 2026 & \textbf{Full posegraph rebuild.}  Re-ran 812\,s mapping
              to completion (84\,494 items).  Result: 701$\times$755
              grid (14.9\,MB data + 41.4\,MB graph), vs.\ prior
              603$\times$266 partial (0.9\,MB data). \\
Mar 2, 2026 & \textbf{Re-ran all benchmarks with full posegraph.}
              dynamics\_0 5-config ablation: L+D achieves APE median
              1.23\,m (73\% improvement over L baseline 4.63\,m).
              Cross-sequence L-only: dynamics\_0 4.63\,m, static\_0
              5.99\,m, lt\_changes\_0 6.32\,m.  Loop-closure jumps
              observed in longer sequences. \\
Mar 2, 2026 & \textbf{Phase 1 LILocBench offline ablation.}  Extracted
              432 frames from dynamics\_0 at 3\,Hz.  DA3-Small live
              inference on GTX 1060.  Baseline+A1--A4 complete:
              A2/A3 IoU 0.783, A4 detection rate 88.3\%.
              A5/A6 deferred (SAM2 not installed locally). \\
Mar 2, 2026 & \textbf{Phase 2 Jetson: marked deferred.}  TRT build
              ready but requires physical RC car platform. \\
            & \textit{(More entries to be added as work progresses)} \\
\bottomrule
\end{longtable}

\end{document}
